---
title: "Aztec Agricultural Productivity Model"
subtitle: "Part 1: Environmental Data"
author: "Rudolf Cesaretti"
date: "Last run on `r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
    number_sections: true
bibliography: References.bib
csl: apa.csl
link-citations: yes
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

```{r, setup, include=FALSE,echo=FALSE, message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=75),tidy=TRUE)
#
rm(list = ls())
```

This R markdown document prepares the data used in the Aztec agricultural productivity random forest model, which is undertaken in Part 2 (the subsequent R markdown script). This involves the following:

  1. Load, reorganize and integrate the modern agricultural analogue dataset, which includes:
      + Contemporary municipio polygons from INEGI
      + 2010 municipal population census statistics
      + 2007 municipal agricultural census statistics
      + FASII and INEGI geospatial land use polygons from the 2000s and 2010s
      + SIAP 2003-2021 crop yield data
      + 
  2. Load, Resample and Calculate topographic variable rasters, which include:
      + 
      + 
      + 
      + 
      + 
      + 
      + 
      + 
      + 
  3. Load, resample and RF impute ISRIC SoilGrids250 rasters, which include:
      + 
      + 
      + 
      + 
      + 
  4. Load and resample CHELSA monthly climate variable rasters, which include:
      + 
      + 
      + 
  
  Topographic/environmental metrics
  
Function to Interpolate Missing Soil Raster Values Using Random Forest Model

Function for RF soil raster imputation of missing values

Resample raster data to 100m x 100m

CRS = WGS84 UTM Zone 14N (EPSG: 32614)


# Setup 

All of the data and scripts are downloadable from the [Aztec Agriculture github repository](https://https://github.com/rcesaret/AztecAgriculture), which can be downloaded locally as a .zip folder or cloned to your own account.

Either way, once you have done so, you will need to modify the working directory (setwd("C:/...)") path and "dir" variables in the code chunk below to match the repository location on your computer.

```{r, label='Set Local Directory Location', message=FALSE,warning=FALSE}

wd <- list()

#SET YOUR LOCAL DIRECTORY LOCATION HERE:
wd$dir <- "D:/Dropbox (ASU)/AztecAgricultureModel/AztecAgriculture/"
#wd$dir <- "C:/Users/TJ McMote/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"

wd$analysis <- paste0(wd$dir,"analysis/")
wd$data_r <- paste0(wd$dir,"data-raw/")
wd$data_p <- paste0(wd$dir,"data-processed/")
wd$data_f <- paste0(wd$dir,"data-final-outputs/")
wd$figs <- paste0(wd$dir,"figures/")
wd$funcs <- paste0(wd$dir,"functions/")

```



## Load R Packages and Custom Functions

```{r, label='Load Libraries', message=FALSE,warning=FALSE}
# Package names
packages <- c("tidyverse", "rgdal", "rgeos", "sp", "sf", "GISTools", "raster", 
              "Matrix", "terra","gdistance", "lwgeom", "tidyr", "stars", "dismo", 
              "purrr", "spatialEco", "whitebox", "classInt","ggnewscale", "lbmech", 
              "data.table", "tidyterra","gridExtra", "cowplot", "scam", "rmarkdown", 
              "spatialreg","spdep", "ggridges", "ggnewscale", "scales", "ggstatsplot",
              "stringi", "fuzzyjoin", "mgcv", "randomForest", "ranger", "kableExtra")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages,installed_packages)

#Read in custom R functions located in the wd$funcs directory folder
FUNCS <- list("RescaleSpatRast.R", "Terra_df.R", "RF_ImputeRast.R")#"splitByAttributes.R", 
invisible(lapply(FUNCS, function(x) source(paste0(wd$funcs,x))))
rm(FUNCS)

```









# Topographic Variables




```{r, label='Load, Resample and Calculate Topographic Variable Rasters', message=FALSE,warning=FALSE}


resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
resample_100m <- disagg(resample_30m, 3)
resample_100m <- aggregate(resample_100m, 10)

## Import INEGI lidar-derived 15-meter DEM
## Continuo de Elevaciones Mexicanas 3.0 (2012)
## https://www.inegi.org.mx/app/geo2/elevacionesmex/#:~:text=El%20Continuo%20de%20Elevaciones%20Mexicano,le%20integran%20valores%20que%20representan


DEM <- rast(paste0(wd$data_r,"DEM_CEM_v3.0_15m_CMex.tif"))

DEM[DEM == 0 | is.nan(DEM)] <- NA

## SMOOTH RASTER
DEM <- focal(DEM, w=5, fun='mean', na.policy="only", na.rm=T)
DEM <- focal(DEM, w=5, fun='mean', na.policy="only", na.rm=T)
DEM <- focal(DEM, w=5, fun='mean', na.policy="only", na.rm=T)

DEM <- focal(DEM, w=5, fun='mean', na.rm=T)
DEM <- focal(DEM, w=7, fun='mean', na.rm=T)
DEM <- focal(DEM, w=9, fun='mean', na.rm=T)
DEM <- focal(DEM, w=11, fun='mean', na.rm=T)
DEM <- focal(DEM, w=13, fun='mean', na.rm=T)

writeRaster(DEM,paste0(wd$data_r,"DEM.tif"), overwrite=TRUE)
#DEM <- rast(paste0(wd$data_r,"DEM.tif"))

DEM_r <- resample(DEM, resample_100m, method = "bilinear")
writeRaster(DEM_r,paste0(wd$data_r,"DEM_r.tif"), overwrite=TRUE)

rm(DEM)

#create temp directory for rasters we will delete later
# we will stack and save these together at the end of this chunk

##### dir.create(paste0(wd$data_p,"temp_rasts"))

######### TRI ######### 

TRI <- terra::terrain(DEM_r, v="TRI")# calc Terrain Ruggedness Index
#TRI_r <- resample(TRI, resample_100m, method = "bilinear")


slope <- terra::terrain(DEM_r, v="slope", unit="degrees")#Slope
#slope_r <- resample(slope, resample_100m, method = "bilinear")
writeRaster(slope,paste0(wd$data_p,"temp_rasts/Slope.tif"), overwrite=TRUE)

######### TWI ######### 

wbt_breach_depressions(dem = paste0(wd$data_r,"DEM_r.tif"), output = paste0(wd$data_p,"temp_rasts/DEM_breach.tif"))
#DEM_breach <- rast(paste0(wd$data_p, "temp_rasts/DEM_breach.tif"))
#DEM_breach <- rast(paste0(wd$data_r,"DEM_breach.tif"))
#writeRaster(DEM_breach,paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), overwrite=TRUE)
#writeRaster(DEM_breach,paste0(wd$data_r,"DEM_breach.tif"), overwrite=TRUE)


wbt_d8_flow_accumulation(paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), output=paste0(wd$data_p,"temp_rasts/Accum.tif"), out_type = "specific contributing area")
Accum <- rast(paste0(wd$data_p, "temp_rasts/Accum.tif"))
#Accum <- rast(paste0(wd$data_r, "Accum.tif"))
#Accum[is.nan(Accum)] <- NA
#Accum_r <- resample(Accum, resample_100m, method = "bilinear")
#writeRaster(Accum,paste0(wd$data_p,"temp_rasts/Accum.tif"), overwrite=TRUE)
#writeRaster(Accum,paste0(wd$data_r,"Accum.tif"), overwrite=TRUE)

wbt_wetness_index(sca=paste0(wd$data_p,"temp_rasts/Accum.tif"), slope=paste0(wd$data_p,"temp_rasts/Slope.tif"), output=paste0(wd$data_p,"temp_rasts/TWI.tif"), verbose_mode = FALSE)
TWI <- rast(paste0(wd$data_p, "temp_rasts/TWI.tif"))
TWI <- focal(TWI, w=17, fun='max', na.policy="only", na.rm=T)
#TWI <- rast(paste0(wd$data_r, "TWI.tif"))
#TWI_r <- resample(TWI, resample_100m, method = "bilinear")
writeRaster(TWI,paste0(wd$data_p,"temp_rasts/TWI.tif"), overwrite=TRUE)
#writeRaster(TWI,paste0(wd$data_r,"TWI.tif"), overwrite=TRUE)


######### SPI ######### 

wbt_stream_power_index(sca=paste0(wd$data_p,"temp_rasts/Accum.tif"),slope=paste0(wd$data_p,"temp_rasts/Slope.tif"),output=paste0(wd$data_p,"temp_rasts/SPI.tif"),exponent = 1)
SPI <- rast(paste0(wd$data_p, "temp_rasts/SPI.tif"))
#SPI <- rast(paste0(wd$data_r, "SPI.tif"))
#SPI_r <- resample(SPI, resample_100m, method = "bilinear")
#writeRaster(SPI,paste0(wd$data_p,"temp_rasts/SPI.tif"), overwrite=TRUE)
#writeRaster(SPI,paste0(wd$data_r,"SPI.tif"), overwrite=TRUE)

######### STI ######### 

wbt_fd8_flow_accumulation(paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), output=paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"), out_type = "specific contributing area")
Accum_fd8 <- rast(paste0(wd$data_p, "temp_rasts/Accum_fd8.tif"))
#Accum_fd8 <- rast(paste0(wd$data_r, "Accum_fd8.tif"))
#writeRaster(Accum_fd8,paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"), overwrite=TRUE)
#writeRaster(Accum_fd8,paste0(wd$data_r,"Accum_fd8.tif"), overwrite=TRUE)

wbt_sediment_transport_index(sca=paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"),slope=paste0(wd$data_p,"temp_rasts/Slope.tif"),output=paste0(wd$data_p,"temp_rasts/STI.tif"))
STI <- rast(paste0(wd$data_p, "temp_rasts/STI.tif"))
#STI <- rast(paste0(wd$data_r, "STI.tif"))
#STI_r <- resample(STI, resample_100m, method = "bilinear")
#writeRaster(STI,paste0(wd$data_p,"temp_rasts/STI.tif"), overwrite=TRUE)
#writeRaster(STI,paste0(wd$data_r,"STI.tif"), overwrite=TRUE)



TopoEnvData <- (c(DEM_r, slope, Accum, Accum_fd8, TRI, TWI, STI, SPI))

rm(DEM_r, slope, Accum, Accum_fd8, TRI, TWI, STI, SPI)

nam = c("DEM", "Slope", "Accum", "Accum_fd8", "TRI", "TWI", "STI", "SPI")

names(TopoEnvData) = nam

writeRaster(TopoEnvData, filename = paste0(wd$data_r, "TopoEnvData.tif"), overwrite=TRUE)

TopoEnv_r <- rast(paste0(wd$data_r, "TopoEnvData.tif"))

rm(TopoEnvData)
```




# Soil Data

ISRIC  SoilGrids250

https://www.isric.org/explore/soilgrids

```{r, label='Soil Depth huge downloads cropping', eval = FALSE}

# # SCRIPT FOR SAVING MULTI-GIGABYTE SOILGRID DOWNLOADS 
# # TO HUEXTLAXANAL2022 DATA-RAW/SOILGRIDS FOLDER

BDRICM <- raster("C:/Users/rcesaret/Downloads/BDRICM_M_250m_ll.tif")
X <- projectRaster(WRB, crs = "+proj=longlat +datum=WGS84 +no_defs")
r2 = raster(extent(X), resolution = c(0.00208, 0.00208), crs = crs(X))
BDRICM <- crop(BDRICM, r2)
BDRICM <- projectRaster(BDRICM, crs = prj)
BDRICM <- resample(BDRICM, BD, method = "bilinear")
writeRaster(BDRICM,paste0(wd$data_r,"SoilGrids/BDRICM.tif"), overwrite=TRUE)


BDRLOG <- raster("C:/Users/rcesaret/Downloads/BDRLOG_M_250m_ll.tif")
X <- projectRaster(WRB, crs = "+proj=longlat +datum=WGS84 +no_defs")
r2 = raster(extent(X), resolution = c(0.00208, 0.00208), crs = crs(X))
BDRLOG <- crop(BDRLOG, r2)
BDRLOG <- projectRaster(BDRLOG, crs = prj)
BDRLOG <- resample(BDRLOG, BD, method = "bilinear")
writeRaster(BDRLOG,paste0(wd$data_r,"SoilGrids/BDRLOG.tif"), overwrite=TRUE)


BDTICM <- raster("C:/Users/rcesaret/Downloads/BDTICM_M_250m_ll.tif")
X <- projectRaster(WRB, crs = "+proj=longlat +datum=WGS84 +no_defs")
r2 = raster(extent(X), resolution = c(0.00208, 0.00208), crs = crs(X))
BDTICM <- crop(BDTICM, r2)
BDTICM <- projectRaster(BDTICM, crs = prj)
BDTICM <- resample(BDTICM, BD, method = "bilinear")
writeRaster(BDTICM,paste0(wd$data_r,"SoilGrids/BDTICM.tif"), overwrite=TRUE)

```



```{r SoilGrids250Vars, echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.pos='!h'}
library(kableExtra)

Variable <- c("BDRICM", "BDRLOG", "BDTICM", "BD", "S", "Z", "C", "SOC", "CEC", "OCD", "N", "SOCS", "WRB")
Description <- c("Depth to bedrock (R horizon) up to 200 cm", "Probability of occurrence of R horizon", "Absolute depth to bedrock", "Bulk density (fine earth fraction)", "Weight percentage of the sand particles (0.05–2 mm)", "Weight percentage of the silt particles (0.0002–0.05 mm)", "Weight percentage of the clay particles (<0.0002 mm)", "Soil organic carbon content", "Cation exchange capacity at ph 7", "Soil organic carbon density", "Nitrogen content", "Soil organic carbon stock", "World Reference Base (2006) Soil Groups")
Units <- c("cm", "percentage", "cm", "kg/m^3", "percentage", "percentage", "percentage", "permille", "	cmolc/kg", "kg/m^3", "cg/kg", "ton/ha", "type (factor)")

data.frame(Variable, Description, Units) %>% 
    kbl(format = 'html',
        escape = FALSE,
        caption = "SoilGrids250 Rasters / Variables") %>% 
    kable_styling(font_size = 22) %>%
    gsub("font-size: initial !important;", 
         "font-size: 22pt !important;", 
         .)

rm(Variable, Description, Units)
```



```{r, label='Load, resample and RF impute Soil Rasters', message=FALSE, warning=FALSE}
###############################################
#resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
#resample_100m <- disagg(resample_30m, 3)
#resample_100m <- aggregate(resample_100m, 10)
#TopoEnv_r <- rast(paste0(wd$data_r, "TopoEnvData.tif"))
###############################################

# Covariates for input into RF_ImputeRast function
DEM_r <- TopoEnv_r[[1]]
Slope <- TopoEnv_r[[2]]

#Municipios2000s <- st_read(paste0(wd$data_r, "AG_Model_2000s/CMex_Municipios_2000s.gpkg"))

# Create Municipios Mask Raster for input into RF_ImputeRast function
Municipios2000s <- st_read(paste0(wd$data_r, "AG_Model_2000s/CMex_Municipios_2000s.gpkg"))
resample_100m_b = resample_100m
values(resample_100m_b) <- 1
m <- mask(resample_100m_b, Municipios2000s)
rm(resample_100m_b,Municipios2000s)

# Depths for each of the first three SoilGrids250 levels 
# for taking the weighted average for 0-30cm
DepthWeights = c(5,10,15) 


# Depth to bedrock (R horizon) up to 200 cm
BDRICM <- rast(paste0(wd$data_r, "SoilGrids/BDRICM.tif"))
BDRICM[BDRICM == 0] <- NA
BDRICM <- focal(BDRICM, w=5, fun='max', na.policy="only", na.rm=T)
BDRICM <- focal(BDRICM, w=5, fun='max', na.policy="only", na.rm=T)
BDRICM <- resample(BDRICM, resample_100m, method = "bilinear")
names(BDRICM) <- "BDRICM"
BDRICMi <- RF_ImputeRast(r = BDRICM, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
BDRICM <- BDRICMi[[1]]
rm(BDRICMi)

#	Probability of occurrence of R horizon	percentage
BDRLOG <- rast(paste0(wd$data_r, "SoilGrids/BDRLOG.tif"))
BDRLOG[BDRLOG == 0] <- NA
BDRLOG <- focal(BDRLOG, w=5, fun='min', na.policy="only", na.rm=T)
BDRLOG <- focal(BDRLOG, w=5, fun='min', na.policy="only", na.rm=T)
BDRLOG <- resample(BDRLOG, resample_100m, method = "bilinear")
names(BDRLOG) <- "BDRLOG"
BDRLOGi <- RF_ImputeRast(r = BDRLOG, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
BDRLOG <- BDRLOGi[[1]]
rm(BDRLOGi)


# Absolute depth to bedrock	cm
BDTICM <- rast(paste0(wd$data_r, "SoilGrids/BDTICM.tif"))
BDTICM[BDTICM == 0] <- NA
BDTICM <- focal(BDTICM, w=5, fun='max', na.policy="only", na.rm=T)
BDTICM <- focal(BDTICM, w=5, fun='max', na.policy="only", na.rm=T)
BDTICM <- resample(BDTICM, resample_100m, method = "bilinear")
names(BDTICM) <- "BDTICM"
BDTICMi <- RF_ImputeRast(r = BDTICM, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
BDTICM <- BDTICMi[[1]]
rm(BDTICMi)

# World Reference Base (2006) Soil Groups
#WRB <- rast(paste0(wd$data_r, "SoilGrids/WRB.tif"))
#WRB <- resample(WRB, resample_100m, method = "bilinear")
#names(WRB) <- "WRB"

# Bulk Density
BD_0_5 <- raster(paste0(wd$data_r, "SoilGrids/BD_0_5.tif"))
BD_0_5[BD_0_5 == 0] <- NA
BD_5_15 <- raster(paste0(wd$data_r, "SoilGrids/BD_5_15.tif"))
BD_5_15[BD_5_15 == 0] <- NA
BD_15_30 <- raster(paste0(wd$data_r, "SoilGrids/BD_15_30.tif"))
BD_15_30[BD_15_30 == 0] <- NA
BD = stack(BD_0_5, BD_5_15, BD_15_30)
BD = weighted.mean(BD, DepthWeights, na.rm=T)
BD <- resample(rast(BD), resample_100m, method = "bilinear")
rm(BD_0_5, BD_5_15, BD_15_30)
names(BD) <- "BD"
BDi <- RF_ImputeRast(r = BD, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
BD <- BDi[[1]]
rm(BDi)

# Clay fraction
C_0_5 <- raster(paste0(wd$data_r, "SoilGrids/C_0_5.tif"))
C_0_5[C_0_5 == 0] <- NA
C_5_15 <- raster(paste0(wd$data_r, "SoilGrids/C_5_15.tif"))
C_5_15[C_5_15 == 0] <- NA
C_15_30 <- raster(paste0(wd$data_r, "SoilGrids/C_15_30.tif"))
C_15_30[C_15_30 == 0] <- NA
C = stack(C_0_5, C_5_15, C_15_30)
C = weighted.mean(C, DepthWeights, na.rm=T)
C <- resample(rast(C), resample_100m, method = "bilinear")
rm(C_0_5, C_5_15, C_15_30)
names(C) <- "C"
Ci <- RF_ImputeRast(r = C, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
C <- Ci[[1]]
rm(Ci)

# Silt fraction
Z_0_5 <- raster(paste0(wd$data_r, "SoilGrids/Z_0_5.tif"))
Z_0_5[Z_0_5 == 0] <- NA
Z_5_15 <- raster(paste0(wd$data_r, "SoilGrids/Z_5_15.tif"))
Z_5_15[Z_5_15 == 0] <- NA
Z_15_30 <- raster(paste0(wd$data_r, "SoilGrids/Z_15_30.tif"))
Z_15_30[Z_15_30 == 0] <- NA
Z = stack(Z_0_5, Z_5_15, Z_15_30)
Z = weighted.mean(Z, DepthWeights, na.rm=T)
Z <- resample(rast(Z), resample_100m, method = "bilinear")
rm(Z_0_5, Z_5_15, Z_15_30)
names(Z) <- "Z"
Zi <- RF_ImputeRast(r = Z, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
Z <- Zi[[1]]
rm(Zi)

# Sand fraction
S_0_5 <- raster(paste0(wd$data_r, "SoilGrids/S_0_5.tif"))
S_0_5[S_0_5 == 0] <- NA
S_5_15 <- raster(paste0(wd$data_r, "SoilGrids/S_5_15.tif"))
S_5_15[S_5_15 == 0] <- NA
S_15_30 <- raster(paste0(wd$data_r, "SoilGrids/S_15_30.tif"))
S_15_30[S_15_30 == 0] <- NA
S = stack(S_0_5, S_5_15, S_15_30)
S = weighted.mean(S, DepthWeights, na.rm=T)
S <- resample(rast(S), resample_100m, method = "bilinear")
rm(S_0_5, S_5_15, S_15_30)
names(S) <- "S"
Si <- RF_ImputeRast(r = S, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
S <- Si[[1]]
rm(Si)

# Soil Organic Carbon content
SOC_0_5 <- raster(paste0(wd$data_r, "SoilGrids/SOC_0_5.tif"))
SOC_0_5[SOC_0_5 == 0] <- NA
SOC_5_15 <- raster(paste0(wd$data_r, "SoilGrids/SOC_5_15.tif"))
SOC_5_15[SOC_5_15 == 0] <- NA
SOC_15_30 <- raster(paste0(wd$data_r, "SoilGrids/SOC_15_30.tif"))
SOC_15_30[SOC_15_30 == 0] <- NA
SOC = stack(SOC_0_5, SOC_5_15, SOC_15_30)
SOC = weighted.mean(SOC, DepthWeights, na.rm=T)
SOC <- resample(rast(SOC), resample_100m, method = "bilinear")
rm(SOC_0_5, SOC_5_15, SOC_15_30)
names(SOC) <- "SOC"
SOCi <- RF_ImputeRast(r = SOC, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
SOC <- SOCi[[1]]
rm(SOCi)

# Cation Exchange Capacity
CEC_0_5 <- raster(paste0(wd$data_r, "SoilGrids/CEC_0_5.tif"))
CEC_0_5[CEC_0_5 == 0] <- NA
CEC_5_15 <- raster(paste0(wd$data_r, "SoilGrids/CEC_5_15.tif"))
CEC_5_15[CEC_5_15 == 0] <- NA
CEC_15_30 <- raster(paste0(wd$data_r, "SoilGrids/CEC_15_30.tif"))
CEC_15_30[CEC_15_30 == 0] <- NA
CEC = stack(CEC_0_5, CEC_5_15, CEC_15_30)
CEC = weighted.mean(CEC, DepthWeights, na.rm=T)
CEC <- resample(rast(CEC), resample_100m, method = "bilinear")
rm(CEC_0_5, CEC_5_15, CEC_15_30)
names(CEC) <- "CEC"
CECi <- RF_ImputeRast(r = CEC, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
CEC <- CECi[[1]]
rm(CECi)

# Organic Carbon Density
OCD_0_5 <- raster(paste0(wd$data_r, "SoilGrids/OCD_0_5.tif"))
OCD_0_5[OCD_0_5 == 0] <- NA
OCD_5_15 <- raster(paste0(wd$data_r, "SoilGrids/OCD_5_15.tif"))
OCD_5_15[OCD_5_15 == 0] <- NA
OCD_15_30 <- raster(paste0(wd$data_r, "SoilGrids/OCD_15_30.tif"))
OCD_15_30[OCD_15_30 == 0] <- NA
OCD = stack(OCD_0_5, OCD_5_15, OCD_15_30)
OCD = weighted.mean(OCD, DepthWeights, na.rm=T)
OCD <- resample(rast(OCD), resample_100m, method = "bilinear")
rm(OCD_0_5, OCD_5_15, OCD_15_30)
names(OCD) <- "OCD"
OCDi <- RF_ImputeRast(r = OCD, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
OCD <- OCDi[[1]]
rm(OCDi)

# Nitrogen
N_0_5 <- raster(paste0(wd$data_r, "SoilGrids/N_0_5.tif"))
N_0_5[N_0_5 == 0] <- NA
N_5_15 <- raster(paste0(wd$data_r, "SoilGrids/N_5_15.tif"))
N_5_15[N_5_15 == 0] <- NA
N_15_30 <- raster(paste0(wd$data_r, "SoilGrids/N_15_30.tif"))
N_15_30[N_15_30 == 0] <- NA
N = stack(N_0_5, N_5_15, N_15_30)
N = weighted.mean(N, DepthWeights, na.rm=T)
N <- resample(rast(N), resample_100m, method = "bilinear")
rm(N_0_5, N_5_15, N_15_30)
names(N) <- "N"
Ni <- RF_ImputeRast(r = N, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
N <- Ni[[1]]
rm(Ni)

#Soil organic carbon stock
SOCS <- rast(paste0(wd$data_r, "SoilGrids/SOCS_0_30.tif"))
SOCS[SOCS == 0] <- NA
SOCS <- resample(SOCS, resample_100m, method = "bilinear")
names(SOCS) <- "SOCS"
SOCSi <- RF_ImputeRast(r = SOCS, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
SOCS <- SOCSi[[1]]
rm(SOCSi)

#Stack Rasters
Soil = c(BDRICM, BDRLOG, BDTICM, BD, S, Z, C, SOC, CEC, OCD, N, SOCS)

# Export Soil Raster Stack
writeRaster(Soil, filename = paste0(wd$data_r, "Soil_r_resampled.tif"), overwrite=TRUE)

#Remove unnecessary variables
rm(DEM_r, Slope, DepthWeights, m)
rm(BDRICM, BDRLOG, BDTICM, BD, S, Z, C, SOC, CEC, OCD, N, SOCS)

```


### Function to Interpolate Missing Soil Raster Values Using Random Forest Model

```{r, label='Function for RF soil raster imputation of missing values', message=FALSE,warning=FALSE, eval = FALSE}

resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
resample_100m <- disagg(resample_30m, 3)
resample_100m <- aggregate(resample_100m, 10)

DEM_r <- TopoEnv_r[[1]]
Slope <- TopoEnv_r[[2]]
Municipios2000s <- st_read(paste0(wd$data_r, "AG_Model_2000s/CMex_Municipios_2000s.gpkg"))

resample_100m_b = resample_100m
values(resample_100m_b) <- 1
m <- mask(resample_100m_b, Municipios2000s)#Data2000s_poly)

r = SOCS
rnam = names(r)
covs = c(DEM_r,Slope)
Mask = m
SampFrac = 0.3
save_mem = FALSE

library(tidyverse)
library(terra)
library(ranger)

RF_ImputeRast <- function(r, # Terra SpatRast raster object
                          rnam = names(r), # name for df variable for raster values from focal SpatRast object
                          covs = NULL, # Terra SpatRast raster object with 1 or more stacked layers 
                          Mask = NULL, # mask of focal area under consideration. Used in reducing the sample for the training dataset; training dataset first excludes areas outside of the mask before applying the sampling fraction
                          SampFrac = 0.3, # sampling fraction for training dataset; applied after mask exclusions, if present (see Mask, above)
                          save_mem = FALSE # whether the save.memory parameter should be used for the ranger random forest function
) {
  
  if (!is.null(Mask)){
    m_df = Terra_df(Mask, "mask")
    names(m_df) <- c("x", "y", "mask", "outside")
    r_df = Terra_df(r, rnam)
    r_df = merge(r_df, m_df, all = F, sort = F)
  } else {
    r_df = Terra_df(r, rnam)
  }
  
  if (!is.null(covs)){
    out_list <- list()  # or result_vector <- c()
    for (i in 1:nlyr(covs)) {
      out_list[[i]] = Terra_df(covs[[i]], names(covs[[i]]))
      out_list[[i]] = out_list[[i]] %>% select(-na_cells)
    }
    covs_df <- Reduce(function(x, y) merge(x, y, all = F, sort = F), out_list)
    r_df = merge(r_df, covs_df, all = F, sort = F)
  }
  
  if (!is.null(Mask)){
    
    train = r_df %>% filter(outside == F) %>% filter(complete.cases(.)) %>% 
      select(-na_cells, -outside, -mask) %>% sample_frac(SampFrac)
    
  } else {
    
    train = r_df %>% filter(complete.cases(.)) %>% select(-na_cells) %>% 
      sample_frac(SampFrac)
  }
  
  f <- names(train)
  f <- f[f != names(r)]
  form <- paste(names(r), "~", paste(f, collapse = " + "))
  form <- as.formula(form)
  
  model <- ranger(
    formula = form,
    data = train,
    save.memory = save_mem 
  )
  
  if (!is.null(Mask)){
    
    pred = r_df %>% select(-na_cells, -outside, -mask) %>% select(-!!sym(rnam)) %>% filter(complete.cases(.))
    
  } else {
    
    pred = r_df %>% select(-na_cells) %>% select(-!!sym(rnam)) %>% filter(complete.cases(.))
    
  }
  
  predicted_vals <- predict(model, pred)
  pred$pred <- predicted_vals$predictions
  
  x = r_df %>% left_join(pred, by = f)
  
  r2 = r
  raster_values <- values(r2)
  missing_values <- is.na(raster_values)
  df_values <- x$pred[missing_values]
  r2[missing_values] <- df_values
  names(r2) <- rnam
  
  out <- list()
  out[[1]] <- r2
  out[[2]] <- model$r.squared
  out[[3]] <- model$prediction.error
  out[[4]] <- model$num.trees
  
  return(out)
}


writeRaster(r2,paste0(wd$data_r,"C_RF_interpolated.tif"), overwrite=TRUE)


```





# Climate Data

```{r, label='Load and Resample Climate Rasters', message=FALSE,warning=FALSE}
chwd = "D:/Dropbox (ASU)/AztecAgricultureModel/GIS_Data/CHELSA/"
prj = "+proj=utm +zone=14 +datum=WGS84 +units=m +no_defs"
t <- raster(paste0(wd$data_r, "TopoEnvData.tif"))
EXT <- projectRaster(t[[1]], crs = "+proj=longlat +datum=WGS84 +no_defs")
###############################################
#resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
#resample_100m <- disagg(resample_30m, 3)
#resample_100m <- aggregate(resample_100m, 10)
###############################################
mo = c(31,28,31,30,31,30,31,31,30,31,30,31)
#divide_layer <- function(x, y) {
#  x / y
#}

# precipitation amount; amount refeers to means mass per unit area. precipitation in the earth's
# atmosphere means precipitation of water in all phases.

prec_01 = raster(paste0(chwd, "prec/CHELSA_prec_01_V1.2_land.tif"))
prec_02 = raster(paste0(chwd, "prec/CHELSA_prec_02_V1.2_land.tif"))
prec_03 = raster(paste0(chwd, "prec/CHELSA_prec_03_V1.2_land.tif"))
prec_04 = raster(paste0(chwd, "prec/CHELSA_prec_04_V1.2_land.tif"))
prec_05 = raster(paste0(chwd, "prec/CHELSA_prec_05_V1.2_land.tif"))
prec_06 = raster(paste0(chwd, "prec/CHELSA_prec_06_V1.2_land.tif"))
prec_07 = raster(paste0(chwd, "prec/CHELSA_prec_07_V1.2_land.tif"))
prec_08 = raster(paste0(chwd, "prec/CHELSA_prec_08_V1.2_land.tif"))
prec_09 = raster(paste0(chwd, "prec/CHELSA_prec_09_V1.2_land.tif"))
prec_10 = raster(paste0(chwd, "prec/CHELSA_prec_10_V1.2_land.tif"))
prec_11 = raster(paste0(chwd, "prec/CHELSA_prec_11_V1.2_land.tif"))
prec_12 = raster(paste0(chwd, "prec/CHELSA_prec_12_V1.2_land.tif"))
prec = stack(prec_01, prec_02, prec_03, prec_04, prec_05, prec_06, prec_07, prec_08, prec_09, prec_10, prec_11, prec_12)
prec <- crop(prec, EXT)
prec <- projectRaster(prec, crs = prj)
for (i in 1:nlayers(prec)) {
  prec[[i]] = prec[[i]]/mo[i]
}
names(prec) <- c("prec_01", "prec_02", "prec_03", "prec_04", "prec_05", "prec_06", "prec_07", "prec_08", "prec_09", "prec_10", "prec_11", "prec_12")
rm(prec_01, prec_02, prec_03, prec_04, prec_05, prec_06, prec_07, prec_08, prec_09, prec_10, prec_11, prec_12)
prec <- rast(prec)
prec_r <- resample(prec, resample_100m, method = "bilinear")
rm(prec)
#writeRaster(prec_r, filename = paste0(wd$data_r, "prec_r.tif"))


# daily mean air temperatures at 2 metres averaged over 1 month
tmean_01 = raster(paste0(chwd, "temp/CHELSA_temp10_01_1979-2013_V1.2_land.tif"))
tmean_02 = raster(paste0(chwd, "temp/CHELSA_temp10_02_1979-2013_V1.2_land.tif"))
tmean_03 = raster(paste0(chwd, "temp/CHELSA_temp10_03_1979-2013_V1.2_land.tif"))
tmean_04 = raster(paste0(chwd, "temp/CHELSA_temp10_04_1979-2013_V1.2_land.tif"))
tmean_05 = raster(paste0(chwd, "temp/CHELSA_temp10_05_1979-2013_V1.2_land.tif"))
tmean_06 = raster(paste0(chwd, "temp/CHELSA_temp10_06_1979-2013_V1.2_land.tif"))
tmean_07 = raster(paste0(chwd, "temp/CHELSA_temp10_07_1979-2013_V1.2_land.tif"))
tmean_08 = raster(paste0(chwd, "temp/CHELSA_temp10_08_1979-2013_V1.2_land.tif"))
tmean_09 = raster(paste0(chwd, "temp/CHELSA_temp10_09_1979-2013_V1.2_land.tif"))
tmean_10 = raster(paste0(chwd, "temp/CHELSA_temp10_10_1979-2013_V1.2_land.tif"))
tmean_11 = raster(paste0(chwd, "temp/CHELSA_temp10_11_1979-2013_V1.2_land.tif"))
tmean_12 = raster(paste0(chwd, "temp/CHELSA_temp10_12_1979-2013_V1.2_land.tif"))
tmean = stack(tmean_01, tmean_02, tmean_03, tmean_04, tmean_05, tmean_06, tmean_07, tmean_08, tmean_09, tmean_10, tmean_11, tmean_12)
tmean <- crop(tmean, EXT)
tmean <- projectRaster(tmean, crs = prj)
names(tmean) <- c("tmean_01", "tmean_02", "tmean_03", "tmean_04", "tmean_05", "tmean_06", "tmean_07", "tmean_08", "tmean_09", "tmean_10", "tmean_11", "tmean_12")
rm(tmean_01, tmean_02, tmean_03, tmean_04, tmean_05, tmean_06, tmean_07, tmean_08, tmean_09, tmean_10, tmean_11, tmean_12)
tmean_r <- rast(tmean)
tmean_r <- resample(tmean_r, resample_100m, method = "bilinear")
rm(tmean)

# daily maximum air temperatures at 2 metres averaged over 1 month
tmax_01 = raster(paste0(chwd, "tmax/CHELSA_tmax10_01_1979-2013_V1.2_land.tif"))
tmax_02 = raster(paste0(chwd, "tmax/CHELSA_tmax10_02_1979-2013_V1.2_land.tif"))
tmax_03 = raster(paste0(chwd, "tmax/CHELSA_tmax10_03_1979-2013_V1.2_land.tif"))
tmax_04 = raster(paste0(chwd, "tmax/CHELSA_tmax10_04_1979-2013_V1.2_land.tif"))
tmax_05 = raster(paste0(chwd, "tmax/CHELSA_tmax10_05_1979-2013_V1.2_land.tif"))
tmax_06 = raster(paste0(chwd, "tmax/CHELSA_tmax10_06_1979-2013_V1.2_land.tif"))
tmax_07 = raster(paste0(chwd, "tmax/CHELSA_tmax10_07_1979-2013_V1.2_land.tif"))
tmax_08 = raster(paste0(chwd, "tmax/CHELSA_tmax10_08_1979-2013_V1.2_land.tif"))
tmax_09 = raster(paste0(chwd, "tmax/CHELSA_tmax10_09_1979-2013_V1.2_land.tif"))
tmax_10 = raster(paste0(chwd, "tmax/CHELSA_tmax10_10_1979-2013_V1.2_land.tif"))
tmax_11 = raster(paste0(chwd, "tmax/CHELSA_tmax10_11_1979-2013_V1.2_land.tif"))
tmax_12 = raster(paste0(chwd, "tmax/CHELSA_tmax10_12_1979-2013_V1.2_land.tif"))
tmax = stack(tmax_01, tmax_02, tmax_03, tmax_04, tmax_05, tmax_06, tmax_07, tmax_08, tmax_09, tmax_10, tmax_11, tmax_12)
tmax <- crop(tmax, EXT)
tmax <- projectRaster(tmax, crs = prj)
names(tmax) <- c("tmax_01", "tmax_02", "tmax_03", "tmax_04", "tmax_05", "tmax_06", "tmax_07", "tmax_08", "tmax_09", "tmax_10", "tmax_11", "tmax_12")
rm(tmax_01, tmax_02, tmax_03, tmax_04, tmax_05, tmax_06, tmax_07, tmax_08, tmax_09, tmax_10, tmax_11, tmax_12)
tmax_r <- rast(tmax)
tmax_r <- resample(tmax_r, resample_100m, method = "bilinear")
rm(tmax)

# daily minimum air temperatures at 2 metres averaged over 1 month
tmin_01 = raster(paste0(chwd, "tmin/CHELSA_tmin10_01_1979-2013_V1.2_land.tif"))
tmin_02 = raster(paste0(chwd, "tmin/CHELSA_tmin10_02_1979-2013_V1.2_land.tif"))
tmin_03 = raster(paste0(chwd, "tmin/CHELSA_tmin10_03_1979-2013_V1.2_land.tif"))
tmin_04 = raster(paste0(chwd, "tmin/CHELSA_tmin10_04_1979-2013_V1.2_land.tif"))
tmin_05 = raster(paste0(chwd, "tmin/CHELSA_tmin10_05_1979-2013_V1.2_land.tif"))
tmin_06 = raster(paste0(chwd, "tmin/CHELSA_tmin10_06_1979-2013_V1.2_land.tif"))
tmin_07 = raster(paste0(chwd, "tmin/CHELSA_tmin10_07_1979-2013_V1.2_land.tif"))
tmin_08 = raster(paste0(chwd, "tmin/CHELSA_tmin10_08_1979-2013_V1.2_land.tif"))
tmin_09 = raster(paste0(chwd, "tmin/CHELSA_tmin10_09_1979-2013_V1.2_land.tif"))
tmin_10 = raster(paste0(chwd, "tmin/CHELSA_tmin10_10_1979-2013_V1.2_land.tif"))
tmin_11 = raster(paste0(chwd, "tmin/CHELSA_tmin10_11_1979-2013_V1.2_land.tif"))
tmin_12 = raster(paste0(chwd, "tmin/CHELSA_tmin10_12_1979-2013_V1.2_land.tif"))
tmin = stack(tmin_01, tmin_02, tmin_03, tmin_04, tmin_05, tmin_06, tmin_07, tmin_08, tmin_09, tmin_10, tmin_11, tmin_12)
tmin <- crop(tmin, EXT)
tmin <- projectRaster(tmin, crs = prj)
names(tmin) <- c("tmin_01", "tmin_02", "tmin_03", "tmin_04", "tmin_05", "tmin_06", "tmin_07", "tmin_08", "tmin_09", "tmin_10", "tmin_11", "tmin_12")
rm(tmin_01, tmin_02, tmin_03, tmin_04, tmin_05, tmin_06, tmin_07, tmin_08, tmin_09, tmin_10, tmin_11, tmin_12)
tmin_r <- rast(tmin)
tmin_r <- resample(tmin_r, resample_100m, method = "bilinear")
rm(tmin)

# Precipitation amount"Amount" means mass per unit area. "Precipitation" in the
# earth's atmosphere means precipitation of water in all phases.

pr_01 = raster(paste0(chwd, "pr/CHELSA_pr_01_1981-2010_V.2.1.tif"))
pr_02 = raster(paste0(chwd, "pr/CHELSA_pr_02_1981-2010_V.2.1.tif"))
pr_03 = raster(paste0(chwd, "pr/CHELSA_pr_03_1981-2010_V.2.1.tif"))
pr_04 = raster(paste0(chwd, "pr/CHELSA_pr_04_1981-2010_V.2.1.tif"))
pr_05 = raster(paste0(chwd, "pr/CHELSA_pr_05_1981-2010_V.2.1.tif"))
pr_06 = raster(paste0(chwd, "pr/CHELSA_pr_06_1981-2010_V.2.1.tif"))
pr_07 = raster(paste0(chwd, "pr/CHELSA_pr_07_1981-2010_V.2.1.tif"))
pr_08 = raster(paste0(chwd, "pr/CHELSA_pr_08_1981-2010_V.2.1.tif"))
pr_09 = raster(paste0(chwd, "pr/CHELSA_pr_09_1981-2010_V.2.1.tif"))
pr_10 = raster(paste0(chwd, "pr/CHELSA_pr_10_1981-2010_V.2.1.tif"))
pr_11 = raster(paste0(chwd, "pr/CHELSA_pr_11_1981-2010_V.2.1.tif"))
pr_12 = raster(paste0(chwd, "pr/CHELSA_pr_12_1981-2010_V.2.1.tif"))
pr = stack(pr_01, pr_02, pr_03, pr_04, pr_05, pr_06, pr_07, pr_08, pr_09, pr_10, pr_11, pr_12)
pr <- crop(pr, EXT)
pr <- projectRaster(pr, crs = prj)
for (i in 1:nlayers(pr)) {
  pr[[i]] = pr[[i]]/mo[i]
}
names(pr) <- c("pr_01", "pr_02", "pr_03", "pr_04", "pr_05", "pr_06", "pr_07", "pr_08", "pr_09", "pr_10", "pr_11", "pr_12")
rm(pr_01, pr_02, pr_03, pr_04, pr_05, pr_06, pr_07, pr_08, pr_09, pr_10, pr_11, pr_12)
pr_r <- rast(pr)
pr_r <- resample(pr_r, resample_100m, method = "bilinear")
###rm(pr)

# Total % cloud cover for each month
clt_01 = raster(paste0(chwd, "clt/CHELSA_clt_01_1981-2010_V.2.1.tif"))
clt_02 = raster(paste0(chwd, "clt/CHELSA_clt_02_1981-2010_V.2.1.tif"))
clt_03 = raster(paste0(chwd, "clt/CHELSA_clt_03_1981-2010_V.2.1.tif"))
clt_04 = raster(paste0(chwd, "clt/CHELSA_clt_04_1981-2010_V.2.1.tif"))
clt_05 = raster(paste0(chwd, "clt/CHELSA_clt_05_1981-2010_V.2.1.tif"))
clt_06 = raster(paste0(chwd, "clt/CHELSA_clt_06_1981-2010_V.2.1.tif"))
clt_07 = raster(paste0(chwd, "clt/CHELSA_clt_07_1981-2010_V.2.1.tif"))
clt_08 = raster(paste0(chwd, "clt/CHELSA_clt_08_1981-2010_V.2.1.tif"))
clt_09 = raster(paste0(chwd, "clt/CHELSA_clt_09_1981-2010_V.2.1.tif"))
clt_10 = raster(paste0(chwd, "clt/CHELSA_clt_10_1981-2010_V.2.1.tif"))
clt_11 = raster(paste0(chwd, "clt/CHELSA_clt_11_1981-2010_V.2.1.tif"))
clt_12 = raster(paste0(chwd, "clt/CHELSA_clt_12_1981-2010_V.2.1.tif"))
clt = stack(clt_01, clt_02, clt_03, clt_04, clt_05, clt_06, clt_07, clt_08, clt_09, clt_10, clt_11, clt_12)
clt <- crop(clt, EXT)
clt <- projectRaster(clt, crs = prj)
# Resample to new resolution
r2 = raster(extent(pr), resolution = c(875, 922), crs = crs(pr))
clt <- resample(clt, r2, method = "bilinear")
names(clt) <- c("clt_01", "clt_02", "clt_03", "clt_04", "clt_05", "clt_06", "clt_07", "clt_08", "clt_09", "clt_10", "clt_11", "clt_12")
rm(clt_01, clt_02, clt_03, clt_04, clt_05, clt_06, clt_07, clt_08, clt_09, clt_10, clt_11, clt_12, r2)
clt_r <- rast(clt)
clt_r <- resample(clt_r, resample_100m, method = "bilinear")
rm(clt, pr)

# Surface downwelling shortwave flux in air
rsds_01 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_01_V.2.1.tif"))
rsds_02 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_02_V.2.1.tif"))
rsds_03 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_03_V.2.1.tif"))
rsds_04 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_04_V.2.1.tif"))
rsds_05 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_05_V.2.1.tif"))
rsds_06 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_06_V.2.1.tif"))
rsds_07 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_07_V.2.1.tif"))
rsds_08 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_08_V.2.1.tif"))
rsds_09 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_09_V.2.1.tif"))
rsds_10 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_10_V.2.1.tif"))
rsds_11 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_11_V.2.1.tif"))
rsds_12 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_12_V.2.1.tif"))
rsds = stack(rsds_01, rsds_02, rsds_03, rsds_04, rsds_05, rsds_06, rsds_07, rsds_08, rsds_09, rsds_10, rsds_11, rsds_12)
rsds <- crop(rsds, EXT)
rsds <- projectRaster(rsds, crs = prj)
names(rsds) <- c("rsds_01", "rsds_02", "rsds_03", "rsds_04", "rsds_05", "rsds_06", "rsds_07", "rsds_08", "rsds_09", "rsds_10", "rsds_11", "rsds_12")
rm(rsds_01, rsds_02, rsds_03, rsds_04, rsds_05, rsds_06, rsds_07, rsds_08, rsds_09, rsds_10, rsds_11, rsds_12)
rsds_r <- rast(rsds)
rsds_r <- resample(rsds_r, resample_100m, method = "bilinear")
rm(rsds)


Clim_r <- c(prec_r,pr_r,tmean_r,tmax_r,tmin_r,clt_r,rsds_r)
#names(Clim_r)
writeRaster(Clim_r, filename = paste0(wd$data_r, "Clim_r_resampled.tif"), overwrite=TRUE)


```








