---
title: "Aztec Agricultural Productivity Model"
subtitle: "Part 1: Environmental Data"
author: "Rudolf Cesaretti"
date: "Last run on `r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
    number_sections: true
bibliography: References.bib
csl: apa.csl
link-citations: yes
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

```{r, setup, include=FALSE,echo=FALSE, message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=75),tidy=TRUE)
#
rm(list = ls())
```

This R markdown document prepares the data used in the Aztec agricultural productivity random forest model, which is undertaken in Part 2 (the subsequent R markdown script). This involves the following:

  1. Load, reorganize and integrate the modern agricultural analogue dataset, which includes:
      + Contemporary municipio polygons from INEGI
      + 2010 municipal population census statistics
      + 2007 municipal agricultural census statistics
      + FASII and INEGI geospatial land use polygons from the 2000s and 2010s
      + SIAP 2003-2021 crop yield data
      + 
  2. Load, Resample and Calculate topographic variable rasters, which include:
      + 
      + 
      + 
      + 
      + 
      + 
      + 
      + 
      + 
  3. Load, resample and RF impute ISRIC SoilGrids250 rasters, which include:
      + 
      + 
      + 
      + 
      + 
  4. Load and resample CHELSA monthly climate variable rasters, which include:
      + 
      + 
      + 
  
  Topographic/environmental metrics
  
Function to Interpolate Missing Soil Raster Values Using Random Forest Model

Function for RF soil raster imputation of missing values

Resample raster data to 100m x 100m

CRS = WGS84 UTM Zone 14N (EPSG: 32614)


# Setup 

All of the data and scripts are downloadable from the [Aztec Agriculture github repository](https://https://github.com/rcesaret/AztecAgriculture), which can be downloaded locally as a .zip folder or cloned to your own account.

Either way, once you have done so, you will need to modify the working directory (setwd("C:/...)") path and "dir" variables in the code chunk below to match the repository location on your computer.

```{r, label='Set Local Directory Location', message=FALSE,warning=FALSE}

wd <- list()

#SET YOUR LOCAL DIRECTORY LOCATION HERE:
wd$dir <- "D:/Dropbox (ASU)/AztecAgricultureModel/AztecAgriculture/"
#wd$dir <- "C:/Users/TJ McMote/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"

wd$analysis <- paste0(wd$dir,"analysis/")
wd$data_r <- paste0(wd$dir,"data-raw/")
wd$data_p <- paste0(wd$dir,"data-processed/")
wd$data_f <- paste0(wd$dir,"data-final-outputs/")
wd$figs <- paste0(wd$dir,"figures/")
wd$funcs <- paste0(wd$dir,"functions/")

```



## Load R Packages and Custom Functions

```{r, label='Load Libraries', message=FALSE,warning=FALSE}
# Package names
packages <- c("tidyverse", "rgdal", "rgeos", "sp", "sf", "GISTools", "raster", 
              "Matrix", "terra","gdistance", "lwgeom", "tidyr", "stars", "dismo", 
              "purrr", "spatialEco", "whitebox", "classInt","ggnewscale", "lbmech", 
              "data.table", "tidyterra","gridExtra", "cowplot", "scam", "rmarkdown", 
              "spatialreg","spdep", "ggridges", "ggnewscale", "scales", "ggstatsplot",
              "stringi", "fuzzyjoin", "mgcv", "randomForest", "ranger", "kableExtra", 
              "Boruta", "mlbench", "caret", "readr")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages,installed_packages)

#Read in custom R functions located in the wd$funcs directory folder
FUNCS <- list("RescaleSpatRast.R", "Terra_df.R", "RF_ImputeRast.R")#"splitByAttributes.R", 
invisible(lapply(FUNCS, function(x) source(paste0(wd$funcs,x))))
rm(FUNCS)

```









# Topographic Variables




```{r, label='Load, Resample and Calculate Topographic Variable Rasters', message=FALSE,warning=FALSE}


resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
resample_100m <- disagg(resample_30m, 3)
resample_100m <- aggregate(resample_100m, 10)

## Import INEGI lidar-derived 15-meter DEM
## Continuo de Elevaciones Mexicanas 3.0 (2012)
## https://www.inegi.org.mx/app/geo2/elevacionesmex/#:~:text=El%20Continuo%20de%20Elevaciones%20Mexicano,le%20integran%20valores%20que%20representan


DEM <- rast(paste0(wd$data_r,"DEM_CEM_v3.0_15m_CMex.tif"))

DEM[DEM == 0 | is.nan(DEM)] <- NA

## SMOOTH RASTER
DEM <- focal(DEM, w=5, fun='mean', na.policy="only", na.rm=T)
DEM <- focal(DEM, w=5, fun='mean', na.policy="only", na.rm=T)
DEM <- focal(DEM, w=5, fun='mean', na.policy="only", na.rm=T)

DEM <- focal(DEM, w=5, fun='mean', na.rm=T)
DEM <- focal(DEM, w=7, fun='mean', na.rm=T)
DEM <- focal(DEM, w=9, fun='mean', na.rm=T)
DEM <- focal(DEM, w=11, fun='mean', na.rm=T)
DEM <- focal(DEM, w=13, fun='mean', na.rm=T)

writeRaster(DEM,paste0(wd$data_r,"DEM.tif"), overwrite=TRUE)
#DEM <- rast(paste0(wd$data_r,"DEM.tif"))

DEM_r <- resample(DEM, resample_100m, method = "bilinear")
writeRaster(DEM_r,paste0(wd$data_r,"DEM_r.tif"), overwrite=TRUE)

rm(DEM)

#create temp directory for rasters we will delete later
# we will stack and save these together at the end of this chunk

##### dir.create(paste0(wd$data_p,"temp_rasts"))

######### TRI ######### 

TRI <- terra::terrain(DEM_r, v="TRI")# calc Terrain Ruggedness Index
#TRI_r <- resample(TRI, resample_100m, method = "bilinear")


slope <- terra::terrain(DEM_r, v="slope", unit="degrees")#Slope
#slope_r <- resample(slope, resample_100m, method = "bilinear")
writeRaster(slope,paste0(wd$data_p,"temp_rasts/Slope.tif"), overwrite=TRUE)

######### TWI ######### 

wbt_breach_depressions(dem = paste0(wd$data_r,"DEM_r.tif"), output = paste0(wd$data_p,"temp_rasts/DEM_breach.tif"))
#DEM_breach <- rast(paste0(wd$data_p, "temp_rasts/DEM_breach.tif"))
#DEM_breach <- rast(paste0(wd$data_r,"DEM_breach.tif"))
#writeRaster(DEM_breach,paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), overwrite=TRUE)
#writeRaster(DEM_breach,paste0(wd$data_r,"DEM_breach.tif"), overwrite=TRUE)


wbt_d8_flow_accumulation(paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), output=paste0(wd$data_p,"temp_rasts/Accum.tif"), out_type = "specific contributing area")
Accum <- rast(paste0(wd$data_p, "temp_rasts/Accum.tif"))
#Accum <- rast(paste0(wd$data_r, "Accum.tif"))
#Accum[is.nan(Accum)] <- NA
#Accum_r <- resample(Accum, resample_100m, method = "bilinear")
#writeRaster(Accum,paste0(wd$data_p,"temp_rasts/Accum.tif"), overwrite=TRUE)
#writeRaster(Accum,paste0(wd$data_r,"Accum.tif"), overwrite=TRUE)

wbt_wetness_index(sca=paste0(wd$data_p,"temp_rasts/Accum.tif"), slope=paste0(wd$data_p,"temp_rasts/Slope.tif"), output=paste0(wd$data_p,"temp_rasts/TWI.tif"), verbose_mode = FALSE)
TWI <- rast(paste0(wd$data_p, "temp_rasts/TWI.tif"))
TWI <- focal(TWI, w=17, fun='max', na.policy="only", na.rm=T)
#TWI <- rast(paste0(wd$data_r, "TWI.tif"))
#TWI_r <- resample(TWI, resample_100m, method = "bilinear")
writeRaster(TWI,paste0(wd$data_p,"temp_rasts/TWI.tif"), overwrite=TRUE)
#writeRaster(TWI,paste0(wd$data_r,"TWI.tif"), overwrite=TRUE)


######### SPI ######### 

wbt_stream_power_index(sca=paste0(wd$data_p,"temp_rasts/Accum.tif"),slope=paste0(wd$data_p,"temp_rasts/Slope.tif"),output=paste0(wd$data_p,"temp_rasts/SPI.tif"),exponent = 1)
SPI <- rast(paste0(wd$data_p, "temp_rasts/SPI.tif"))
#SPI <- rast(paste0(wd$data_r, "SPI.tif"))
#SPI_r <- resample(SPI, resample_100m, method = "bilinear")
#writeRaster(SPI,paste0(wd$data_p,"temp_rasts/SPI.tif"), overwrite=TRUE)
#writeRaster(SPI,paste0(wd$data_r,"SPI.tif"), overwrite=TRUE)

######### STI ######### 

wbt_fd8_flow_accumulation(paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), output=paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"), out_type = "specific contributing area")
Accum_fd8 <- rast(paste0(wd$data_p, "temp_rasts/Accum_fd8.tif"))
#Accum_fd8 <- rast(paste0(wd$data_r, "Accum_fd8.tif"))
#writeRaster(Accum_fd8,paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"), overwrite=TRUE)
#writeRaster(Accum_fd8,paste0(wd$data_r,"Accum_fd8.tif"), overwrite=TRUE)

wbt_sediment_transport_index(sca=paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"),slope=paste0(wd$data_p,"temp_rasts/Slope.tif"),output=paste0(wd$data_p,"temp_rasts/STI.tif"))
STI <- rast(paste0(wd$data_p, "temp_rasts/STI.tif"))
#STI <- rast(paste0(wd$data_r, "STI.tif"))
#STI_r <- resample(STI, resample_100m, method = "bilinear")
#writeRaster(STI,paste0(wd$data_p,"temp_rasts/STI.tif"), overwrite=TRUE)
#writeRaster(STI,paste0(wd$data_r,"STI.tif"), overwrite=TRUE)



TopoEnvData <- (c(DEM_r, slope, Accum, Accum_fd8, TRI, TWI, STI, SPI))

rm(DEM_r, slope, Accum, Accum_fd8, TRI, TWI, STI, SPI)

nam = c("DEM", "Slope", "Accum", "Accum_fd8", "TRI", "TWI", "STI", "SPI")

names(TopoEnvData) = nam

writeRaster(TopoEnvData, filename = paste0(wd$data_p, "TopoEnvData.tif"), overwrite=TRUE)

TopoEnv_r <- rast(paste0(wd$data_r, "TopoEnvData.tif"))

rm(TopoEnvData)
```




# Soil Data

ISRIC  SoilGrids250

https://www.isric.org/explore/soilgrids

```{r, label='Soil Depth huge downloads cropping', eval = FALSE}

# # SCRIPT FOR SAVING MULTI-GIGABYTE SOILGRID DOWNLOADS 
# # TO HUEXTLAXANAL2022 DATA-RAW/SOILGRIDS FOLDER

BDRICM <- raster("C:/Users/rcesaret/Downloads/BDRICM_M_250m_ll.tif")
X <- projectRaster(WRB, crs = "+proj=longlat +datum=WGS84 +no_defs")
r2 = raster(extent(X), resolution = c(0.00208, 0.00208), crs = crs(X))
BDRICM <- crop(BDRICM, r2)
BDRICM <- projectRaster(BDRICM, crs = prj)
BDRICM <- resample(BDRICM, BD, method = "bilinear")
writeRaster(BDRICM,paste0(wd$data_r,"SoilGrids/BDRICM.tif"), overwrite=TRUE)


BDRLOG <- raster("C:/Users/rcesaret/Downloads/BDRLOG_M_250m_ll.tif")
X <- projectRaster(WRB, crs = "+proj=longlat +datum=WGS84 +no_defs")
r2 = raster(extent(X), resolution = c(0.00208, 0.00208), crs = crs(X))
BDRLOG <- crop(BDRLOG, r2)
BDRLOG <- projectRaster(BDRLOG, crs = prj)
BDRLOG <- resample(BDRLOG, BD, method = "bilinear")
writeRaster(BDRLOG,paste0(wd$data_r,"SoilGrids/BDRLOG.tif"), overwrite=TRUE)


BDTICM <- raster("C:/Users/rcesaret/Downloads/BDTICM_M_250m_ll.tif")
X <- projectRaster(WRB, crs = "+proj=longlat +datum=WGS84 +no_defs")
r2 = raster(extent(X), resolution = c(0.00208, 0.00208), crs = crs(X))
BDTICM <- crop(BDTICM, r2)
BDTICM <- projectRaster(BDTICM, crs = prj)
BDTICM <- resample(BDTICM, BD, method = "bilinear")
writeRaster(BDTICM,paste0(wd$data_r,"SoilGrids/BDTICM.tif"), overwrite=TRUE)

```



```{r SoilGrids250Vars, echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.pos='!h'}
library(kableExtra)

Variable <- c("BDRICM", "BDRLOG", "BDTICM", "BD", "S", "Z", "C", "SOC", "CEC", "OCD", "N", "SOCS", "WRB")
Description <- c("Depth to bedrock (R horizon) up to 200 cm", "Probability of occurrence of R horizon", "Absolute depth to bedrock", "Bulk density (fine earth fraction)", "Weight percentage of the sand particles (0.05–2 mm)", "Weight percentage of the silt particles (0.0002–0.05 mm)", "Weight percentage of the clay particles (<0.0002 mm)", "Soil organic carbon content", "Cation exchange capacity at ph 7", "Soil organic carbon density", "Nitrogen content", "Soil organic carbon stock", "World Reference Base (2006) Soil Groups")
Units <- c("cm", "percentage", "cm", "kg/m^3", "percentage", "percentage", "percentage", "permille", "	cmolc/kg", "kg/m^3", "cg/kg", "ton/ha", "type (factor)")

data.frame(Variable, Description, Units) %>% 
    kbl(format = 'html',
        escape = FALSE,
        caption = "SoilGrids250 Rasters / Variables") %>% 
    kable_styling(font_size = 22) %>%
    gsub("font-size: initial !important;", 
         "font-size: 22pt !important;", 
         .)

rm(Variable, Description, Units)
```



```{r, label='Load, resample and RF impute Soil Rasters', message=FALSE, warning=FALSE}
###############################################
#resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
#resample_100m <- disagg(resample_30m, 3)
#resample_100m <- aggregate(resample_100m, 10)
#TopoEnv_r <- rast(paste0(wd$data_r, "TopoEnvData.tif"))
###############################################

# Covariates for input into RF_ImputeRast function
DEM_r <- TopoEnv_r[[1]]
Slope <- TopoEnv_r[[2]]

#Municipios2000s <- st_read(paste0(wd$data_r, "AG_Model_2000s/CMex_Municipios_2000s.gpkg"))

# Create Municipios Mask Raster for input into RF_ImputeRast function
Municipios2000s <- st_read(paste0(wd$data_r, "AG_Model_2000s/CMex_Municipios_2000s.gpkg"))
resample_100m_b = resample_100m
values(resample_100m_b) <- 1
m <- mask(resample_100m_b, Municipios2000s)
rm(resample_100m_b,Municipios2000s)

# Depths for each of the first three SoilGrids250 levels 
# for taking the weighted average for 0-30cm
DepthWeights = c(5,10,15) 


# Depth to bedrock (R horizon) up to 200 cm
BDRICM <- rast(paste0(wd$data_r, "SoilGrids/BDRICM.tif"))
BDRICM[BDRICM == 0] <- NA
BDRICM <- focal(BDRICM, w=5, fun='max', na.policy="only", na.rm=T)
BDRICM <- focal(BDRICM, w=5, fun='max', na.policy="only", na.rm=T)
BDRICM <- resample(BDRICM, resample_100m, method = "bilinear")
names(BDRICM) <- "BDRICM"
BDRICMi <- RF_ImputeRast(r = BDRICM, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
BDRICM <- BDRICMi[[1]]
rm(BDRICMi)

#	Probability of occurrence of R horizon	percentage
BDRLOG <- rast(paste0(wd$data_r, "SoilGrids/BDRLOG.tif"))
BDRLOG[BDRLOG == 0] <- NA
BDRLOG <- focal(BDRLOG, w=5, fun='min', na.policy="only", na.rm=T)
BDRLOG <- focal(BDRLOG, w=5, fun='min', na.policy="only", na.rm=T)
BDRLOG <- resample(BDRLOG, resample_100m, method = "bilinear")
names(BDRLOG) <- "BDRLOG"
BDRLOGi <- RF_ImputeRast(r = BDRLOG, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
BDRLOG <- BDRLOGi[[1]]
rm(BDRLOGi)


# Absolute depth to bedrock	cm
BDTICM <- rast(paste0(wd$data_r, "SoilGrids/BDTICM.tif"))
BDTICM[BDTICM == 0] <- NA
BDTICM <- focal(BDTICM, w=5, fun='max', na.policy="only", na.rm=T)
BDTICM <- focal(BDTICM, w=5, fun='max', na.policy="only", na.rm=T)
BDTICM <- resample(BDTICM, resample_100m, method = "bilinear")
names(BDTICM) <- "BDTICM"
BDTICMi <- RF_ImputeRast(r = BDTICM, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
BDTICM <- BDTICMi[[1]]
rm(BDTICMi)

# World Reference Base (2006) Soil Groups
WRB <- rast(paste0(wd$data_r, "SoilGrids/WRB.tif"))
WRB <- as.factor(WRB)
cls = levels(WRB)[[1]] %>% select(-WRB) %>%
  mutate(WRB = case_when(
    ID == 0 ~ "NoData",
    ID == 3 ~ "Andosols",
    ID == 4 ~ "Arenosols",
    ID == 5 ~ "Calcisols",
    ID == 6 ~ "Cambisols",
    ID == 9 ~ "Durisols",
    ID == 11 ~ "Fluvisols",
    ID == 12 ~ "Gleysols",
    ID == 16 ~ "Leptosols",
    ID == 18 ~ "Luvisols",
    ID == 20 ~ "Phaeozems",
    ID == 24 ~ "Regosols",
    ID == 25 ~ "Solonchaks",
    ID == 26 ~ "Solonetz",
    ID == 28 ~ "Umbrisols",
    ID == 29 ~ "Vertisols"))
names(WRB) <- "WRB"
levels(WRB) <- cls
WRB <- resample(WRB, resample_100m, method = "near")

# Bulk Density
BD_0_5 <- raster(paste0(wd$data_r, "SoilGrids/BD_0_5.tif"))
BD_0_5[BD_0_5 == 0] <- NA
BD_5_15 <- raster(paste0(wd$data_r, "SoilGrids/BD_5_15.tif"))
BD_5_15[BD_5_15 == 0] <- NA
BD_15_30 <- raster(paste0(wd$data_r, "SoilGrids/BD_15_30.tif"))
BD_15_30[BD_15_30 == 0] <- NA
BD = stack(BD_0_5, BD_5_15, BD_15_30)
BD = weighted.mean(BD, DepthWeights, na.rm=T)
BD <- resample(rast(BD), resample_100m, method = "bilinear")
rm(BD_0_5, BD_5_15, BD_15_30)
names(BD) <- "BD"
BDi <- RF_ImputeRast(r = BD, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
BD <- BDi[[1]]
rm(BDi)

# Clay fraction
C_0_5 <- raster(paste0(wd$data_r, "SoilGrids/C_0_5.tif"))
C_0_5[C_0_5 == 0] <- NA
C_5_15 <- raster(paste0(wd$data_r, "SoilGrids/C_5_15.tif"))
C_5_15[C_5_15 == 0] <- NA
C_15_30 <- raster(paste0(wd$data_r, "SoilGrids/C_15_30.tif"))
C_15_30[C_15_30 == 0] <- NA
C = stack(C_0_5, C_5_15, C_15_30)
C = weighted.mean(C, DepthWeights, na.rm=T)
C <- resample(rast(C), resample_100m, method = "bilinear")
rm(C_0_5, C_5_15, C_15_30)
names(C) <- "C"
Ci <- RF_ImputeRast(r = C, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
C <- Ci[[1]]
rm(Ci)

# Silt fraction
Z_0_5 <- raster(paste0(wd$data_r, "SoilGrids/Z_0_5.tif"))
Z_0_5[Z_0_5 == 0] <- NA
Z_5_15 <- raster(paste0(wd$data_r, "SoilGrids/Z_5_15.tif"))
Z_5_15[Z_5_15 == 0] <- NA
Z_15_30 <- raster(paste0(wd$data_r, "SoilGrids/Z_15_30.tif"))
Z_15_30[Z_15_30 == 0] <- NA
Z = stack(Z_0_5, Z_5_15, Z_15_30)
Z = weighted.mean(Z, DepthWeights, na.rm=T)
Z <- resample(rast(Z), resample_100m, method = "bilinear")
rm(Z_0_5, Z_5_15, Z_15_30)
names(Z) <- "Z"
Zi <- RF_ImputeRast(r = Z, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
Z <- Zi[[1]]
rm(Zi)

# Sand fraction
S_0_5 <- raster(paste0(wd$data_r, "SoilGrids/S_0_5.tif"))
S_0_5[S_0_5 == 0] <- NA
S_5_15 <- raster(paste0(wd$data_r, "SoilGrids/S_5_15.tif"))
S_5_15[S_5_15 == 0] <- NA
S_15_30 <- raster(paste0(wd$data_r, "SoilGrids/S_15_30.tif"))
S_15_30[S_15_30 == 0] <- NA
S = stack(S_0_5, S_5_15, S_15_30)
S = weighted.mean(S, DepthWeights, na.rm=T)
S <- resample(rast(S), resample_100m, method = "bilinear")
rm(S_0_5, S_5_15, S_15_30)
names(S) <- "S"
Si <- RF_ImputeRast(r = S, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
S <- Si[[1]]
rm(Si)

# Soil Organic Carbon content
SOC_0_5 <- raster(paste0(wd$data_r, "SoilGrids/SOC_0_5.tif"))
SOC_0_5[SOC_0_5 == 0] <- NA
SOC_5_15 <- raster(paste0(wd$data_r, "SoilGrids/SOC_5_15.tif"))
SOC_5_15[SOC_5_15 == 0] <- NA
SOC_15_30 <- raster(paste0(wd$data_r, "SoilGrids/SOC_15_30.tif"))
SOC_15_30[SOC_15_30 == 0] <- NA
SOC = stack(SOC_0_5, SOC_5_15, SOC_15_30)
SOC = weighted.mean(SOC, DepthWeights, na.rm=T)
SOC <- resample(rast(SOC), resample_100m, method = "bilinear")
rm(SOC_0_5, SOC_5_15, SOC_15_30)
names(SOC) <- "SOC"
SOCi <- RF_ImputeRast(r = SOC, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
SOC <- SOCi[[1]]
rm(SOCi)

# Cation Exchange Capacity
CEC_0_5 <- raster(paste0(wd$data_r, "SoilGrids/CEC_0_5.tif"))
CEC_0_5[CEC_0_5 == 0] <- NA
CEC_5_15 <- raster(paste0(wd$data_r, "SoilGrids/CEC_5_15.tif"))
CEC_5_15[CEC_5_15 == 0] <- NA
CEC_15_30 <- raster(paste0(wd$data_r, "SoilGrids/CEC_15_30.tif"))
CEC_15_30[CEC_15_30 == 0] <- NA
CEC = stack(CEC_0_5, CEC_5_15, CEC_15_30)
CEC = weighted.mean(CEC, DepthWeights, na.rm=T)
CEC <- resample(rast(CEC), resample_100m, method = "bilinear")
rm(CEC_0_5, CEC_5_15, CEC_15_30)
names(CEC) <- "CEC"
CECi <- RF_ImputeRast(r = CEC, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
CEC <- CECi[[1]]
rm(CECi)

# Organic Carbon Density
OCD_0_5 <- raster(paste0(wd$data_r, "SoilGrids/OCD_0_5.tif"))
OCD_0_5[OCD_0_5 == 0] <- NA
OCD_5_15 <- raster(paste0(wd$data_r, "SoilGrids/OCD_5_15.tif"))
OCD_5_15[OCD_5_15 == 0] <- NA
OCD_15_30 <- raster(paste0(wd$data_r, "SoilGrids/OCD_15_30.tif"))
OCD_15_30[OCD_15_30 == 0] <- NA
OCD = stack(OCD_0_5, OCD_5_15, OCD_15_30)
OCD = weighted.mean(OCD, DepthWeights, na.rm=T)
OCD <- resample(rast(OCD), resample_100m, method = "bilinear")
rm(OCD_0_5, OCD_5_15, OCD_15_30)
names(OCD) <- "OCD"
OCDi <- RF_ImputeRast(r = OCD, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
OCD <- OCDi[[1]]
rm(OCDi)

# Nitrogen
N_0_5 <- raster(paste0(wd$data_r, "SoilGrids/N_0_5.tif"))
N_0_5[N_0_5 == 0] <- NA
N_5_15 <- raster(paste0(wd$data_r, "SoilGrids/N_5_15.tif"))
N_5_15[N_5_15 == 0] <- NA
N_15_30 <- raster(paste0(wd$data_r, "SoilGrids/N_15_30.tif"))
N_15_30[N_15_30 == 0] <- NA
N = stack(N_0_5, N_5_15, N_15_30)
N = weighted.mean(N, DepthWeights, na.rm=T)
N <- resample(rast(N), resample_100m, method = "bilinear")
rm(N_0_5, N_5_15, N_15_30)
names(N) <- "N"
Ni <- RF_ImputeRast(r = N, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
N <- Ni[[1]]
rm(Ni)

#Soil organic carbon stock
SOCS <- rast(paste0(wd$data_r, "SoilGrids/SOCS_0_30.tif"))
SOCS[SOCS == 0] <- NA
SOCS <- resample(SOCS, resample_100m, method = "bilinear")
names(SOCS) <- "SOCS"
SOCSi <- RF_ImputeRast(r = SOCS, covs = c(DEM_r,Slope), Mask = m, SampFrac = 0.3, save_mem = FALSE)
SOCS <- SOCSi[[1]]
rm(SOCSi)

#Stack Rasters
Soil = c(BDRICM, BDRLOG, BDTICM, BD, S, Z, C, SOC, CEC, OCD, N, SOCS, WRB)

# Export Soil Raster Stack
writeRaster(Soil, filename = paste0(wd$data_p, "Soil_r_resampled.tif"), overwrite=TRUE, append=FALSE)

#Remove unnecessary variables
rm(DEM_r, Slope, DepthWeights, m)
rm(BDRICM, BDRLOG, BDTICM, BD, S, Z, C, SOC, CEC, OCD, N, SOCS, WRB)

```


### Function to Interpolate Missing Soil Raster Values Using Random Forest Model

```{r, label='Function for RF soil raster imputation of missing values', message=FALSE,warning=FALSE, eval = FALSE}

resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
resample_100m <- disagg(resample_30m, 3)
resample_100m <- aggregate(resample_100m, 10)

DEM_r <- TopoEnv_r[[1]]
Slope <- TopoEnv_r[[2]]
Municipios2000s <- st_read(paste0(wd$data_r, "AG_Model_2000s/CMex_Municipios_2000s.gpkg"))

resample_100m_b = resample_100m
values(resample_100m_b) <- 1
m <- mask(resample_100m_b, Municipios2000s)#Data2000s_poly)

r = SOCS
rnam = names(r)
covs = c(DEM_r,Slope)
Mask = m
SampFrac = 0.3
save_mem = FALSE

library(tidyverse)
library(terra)
library(ranger)

RF_ImputeRast <- function(r, # Terra SpatRast raster object
                          rnam = names(r), # name for df variable for raster values from focal SpatRast object
                          covs = NULL, # Terra SpatRast raster object with 1 or more stacked layers 
                          Mask = NULL, # mask of focal area under consideration. Used in reducing the sample for the training dataset; training dataset first excludes areas outside of the mask before applying the sampling fraction
                          SampFrac = 0.3, # sampling fraction for training dataset; applied after mask exclusions, if present (see Mask, above)
                          save_mem = FALSE # whether the save.memory parameter should be used for the ranger random forest function
) {
  
  if (!is.null(Mask)){
    m_df = Terra_df(Mask, "mask")
    names(m_df) <- c("x", "y", "mask", "outside")
    r_df = Terra_df(r, rnam)
    r_df = merge(r_df, m_df, all = F, sort = F)
  } else {
    r_df = Terra_df(r, rnam)
  }
  
  if (!is.null(covs)){
    out_list <- list()  # or result_vector <- c()
    for (i in 1:nlyr(covs)) {
      out_list[[i]] = Terra_df(covs[[i]], names(covs[[i]]))
      out_list[[i]] = out_list[[i]] %>% select(-na_cells)
    }
    covs_df <- Reduce(function(x, y) merge(x, y, all = F, sort = F), out_list)
    r_df = merge(r_df, covs_df, all = F, sort = F)
  }
  
  if (!is.null(Mask)){
    
    train = r_df %>% filter(outside == F) %>% filter(complete.cases(.)) %>% 
      select(-na_cells, -outside, -mask) %>% sample_frac(SampFrac)
    
  } else {
    
    train = r_df %>% filter(complete.cases(.)) %>% select(-na_cells) %>% 
      sample_frac(SampFrac)
  }
  
  f <- names(train)
  f <- f[f != names(r)]
  form <- paste(names(r), "~", paste(f, collapse = " + "))
  form <- as.formula(form)
  
  model <- ranger(
    formula = form,
    data = train,
    save.memory = save_mem 
  )
  
  if (!is.null(Mask)){
    
    pred = r_df %>% select(-na_cells, -outside, -mask) %>% select(-!!sym(rnam)) %>% filter(complete.cases(.))
    
  } else {
    
    pred = r_df %>% select(-na_cells) %>% select(-!!sym(rnam)) %>% filter(complete.cases(.))
    
  }
  
  predicted_vals <- predict(model, pred)
  pred$pred <- predicted_vals$predictions
  
  x = r_df %>% left_join(pred, by = f)
  
  r2 = r
  raster_values <- values(r2)
  missing_values <- is.na(raster_values)
  df_values <- x$pred[missing_values]
  r2[missing_values] <- df_values
  names(r2) <- rnam
  
  out <- list()
  out[[1]] <- r2
  out[[2]] <- model$r.squared
  out[[3]] <- model$prediction.error
  out[[4]] <- model$num.trees
  
  return(out)
}


writeRaster(r2,paste0(wd$data_r,"C_RF_interpolated.tif"), overwrite=TRUE)


```





# Climate Data

```{r, label='Load and Resample Climate Rasters', message=FALSE,warning=FALSE}
chwd = "D:/Dropbox (ASU)/AztecAgricultureModel/GIS_Data/CHELSA/"
prj = "+proj=utm +zone=14 +datum=WGS84 +units=m +no_defs"
t <- raster(paste0(wd$data_r, "TopoEnvData.tif"))
EXT <- projectRaster(t[[1]], crs = "+proj=longlat +datum=WGS84 +no_defs")
###############################################
#resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
#resample_100m <- disagg(resample_30m, 3)
#resample_100m <- aggregate(resample_100m, 10)
###############################################
mo = c(31,28,31,30,31,30,31,31,30,31,30,31)
#divide_layer <- function(x, y) {
#  x / y
#}

# precipitation amount; amount refeers to means mass per unit area. precipitation in the earth's
# atmosphere means precipitation of water in all phases.

prec_01 = raster(paste0(chwd, "prec/CHELSA_prec_01_V1.2_land.tif"))
prec_02 = raster(paste0(chwd, "prec/CHELSA_prec_02_V1.2_land.tif"))
prec_03 = raster(paste0(chwd, "prec/CHELSA_prec_03_V1.2_land.tif"))
prec_04 = raster(paste0(chwd, "prec/CHELSA_prec_04_V1.2_land.tif"))
prec_05 = raster(paste0(chwd, "prec/CHELSA_prec_05_V1.2_land.tif"))
prec_06 = raster(paste0(chwd, "prec/CHELSA_prec_06_V1.2_land.tif"))
prec_07 = raster(paste0(chwd, "prec/CHELSA_prec_07_V1.2_land.tif"))
prec_08 = raster(paste0(chwd, "prec/CHELSA_prec_08_V1.2_land.tif"))
prec_09 = raster(paste0(chwd, "prec/CHELSA_prec_09_V1.2_land.tif"))
prec_10 = raster(paste0(chwd, "prec/CHELSA_prec_10_V1.2_land.tif"))
prec_11 = raster(paste0(chwd, "prec/CHELSA_prec_11_V1.2_land.tif"))
prec_12 = raster(paste0(chwd, "prec/CHELSA_prec_12_V1.2_land.tif"))
prec = stack(prec_01, prec_02, prec_03, prec_04, prec_05, prec_06, prec_07, prec_08, prec_09, prec_10, prec_11, prec_12)
prec <- crop(prec, EXT)
prec <- projectRaster(prec, crs = prj)
for (i in 1:nlayers(prec)) {
  prec[[i]] = prec[[i]]/mo[i]
}
names(prec) <- c("prec_01", "prec_02", "prec_03", "prec_04", "prec_05", "prec_06", "prec_07", "prec_08", "prec_09", "prec_10", "prec_11", "prec_12")
rm(prec_01, prec_02, prec_03, prec_04, prec_05, prec_06, prec_07, prec_08, prec_09, prec_10, prec_11, prec_12)
prec <- rast(prec)
prec_r <- resample(prec, resample_100m, method = "bilinear")
rm(prec)
#writeRaster(prec_r, filename = paste0(wd$data_r, "prec_r.tif"))


# daily mean air temperatures at 2 metres averaged over 1 month
tmean_01 = raster(paste0(chwd, "temp/CHELSA_temp10_01_1979-2013_V1.2_land.tif"))
tmean_02 = raster(paste0(chwd, "temp/CHELSA_temp10_02_1979-2013_V1.2_land.tif"))
tmean_03 = raster(paste0(chwd, "temp/CHELSA_temp10_03_1979-2013_V1.2_land.tif"))
tmean_04 = raster(paste0(chwd, "temp/CHELSA_temp10_04_1979-2013_V1.2_land.tif"))
tmean_05 = raster(paste0(chwd, "temp/CHELSA_temp10_05_1979-2013_V1.2_land.tif"))
tmean_06 = raster(paste0(chwd, "temp/CHELSA_temp10_06_1979-2013_V1.2_land.tif"))
tmean_07 = raster(paste0(chwd, "temp/CHELSA_temp10_07_1979-2013_V1.2_land.tif"))
tmean_08 = raster(paste0(chwd, "temp/CHELSA_temp10_08_1979-2013_V1.2_land.tif"))
tmean_09 = raster(paste0(chwd, "temp/CHELSA_temp10_09_1979-2013_V1.2_land.tif"))
tmean_10 = raster(paste0(chwd, "temp/CHELSA_temp10_10_1979-2013_V1.2_land.tif"))
tmean_11 = raster(paste0(chwd, "temp/CHELSA_temp10_11_1979-2013_V1.2_land.tif"))
tmean_12 = raster(paste0(chwd, "temp/CHELSA_temp10_12_1979-2013_V1.2_land.tif"))
tmean = stack(tmean_01, tmean_02, tmean_03, tmean_04, tmean_05, tmean_06, tmean_07, tmean_08, tmean_09, tmean_10, tmean_11, tmean_12)
tmean <- crop(tmean, EXT)
tmean <- projectRaster(tmean, crs = prj)
names(tmean) <- c("tmean_01", "tmean_02", "tmean_03", "tmean_04", "tmean_05", "tmean_06", "tmean_07", "tmean_08", "tmean_09", "tmean_10", "tmean_11", "tmean_12")
rm(tmean_01, tmean_02, tmean_03, tmean_04, tmean_05, tmean_06, tmean_07, tmean_08, tmean_09, tmean_10, tmean_11, tmean_12)
tmean_r <- rast(tmean)
tmean_r <- resample(tmean_r, resample_100m, method = "bilinear")
rm(tmean)

# daily maximum air temperatures at 2 metres averaged over 1 month
tmax_01 = raster(paste0(chwd, "tmax/CHELSA_tmax10_01_1979-2013_V1.2_land.tif"))
tmax_02 = raster(paste0(chwd, "tmax/CHELSA_tmax10_02_1979-2013_V1.2_land.tif"))
tmax_03 = raster(paste0(chwd, "tmax/CHELSA_tmax10_03_1979-2013_V1.2_land.tif"))
tmax_04 = raster(paste0(chwd, "tmax/CHELSA_tmax10_04_1979-2013_V1.2_land.tif"))
tmax_05 = raster(paste0(chwd, "tmax/CHELSA_tmax10_05_1979-2013_V1.2_land.tif"))
tmax_06 = raster(paste0(chwd, "tmax/CHELSA_tmax10_06_1979-2013_V1.2_land.tif"))
tmax_07 = raster(paste0(chwd, "tmax/CHELSA_tmax10_07_1979-2013_V1.2_land.tif"))
tmax_08 = raster(paste0(chwd, "tmax/CHELSA_tmax10_08_1979-2013_V1.2_land.tif"))
tmax_09 = raster(paste0(chwd, "tmax/CHELSA_tmax10_09_1979-2013_V1.2_land.tif"))
tmax_10 = raster(paste0(chwd, "tmax/CHELSA_tmax10_10_1979-2013_V1.2_land.tif"))
tmax_11 = raster(paste0(chwd, "tmax/CHELSA_tmax10_11_1979-2013_V1.2_land.tif"))
tmax_12 = raster(paste0(chwd, "tmax/CHELSA_tmax10_12_1979-2013_V1.2_land.tif"))
tmax = stack(tmax_01, tmax_02, tmax_03, tmax_04, tmax_05, tmax_06, tmax_07, tmax_08, tmax_09, tmax_10, tmax_11, tmax_12)
tmax <- crop(tmax, EXT)
tmax <- projectRaster(tmax, crs = prj)
names(tmax) <- c("tmax_01", "tmax_02", "tmax_03", "tmax_04", "tmax_05", "tmax_06", "tmax_07", "tmax_08", "tmax_09", "tmax_10", "tmax_11", "tmax_12")
rm(tmax_01, tmax_02, tmax_03, tmax_04, tmax_05, tmax_06, tmax_07, tmax_08, tmax_09, tmax_10, tmax_11, tmax_12)
tmax_r <- rast(tmax)
tmax_r <- resample(tmax_r, resample_100m, method = "bilinear")
rm(tmax)

# daily minimum air temperatures at 2 metres averaged over 1 month
tmin_01 = raster(paste0(chwd, "tmin/CHELSA_tmin10_01_1979-2013_V1.2_land.tif"))
tmin_02 = raster(paste0(chwd, "tmin/CHELSA_tmin10_02_1979-2013_V1.2_land.tif"))
tmin_03 = raster(paste0(chwd, "tmin/CHELSA_tmin10_03_1979-2013_V1.2_land.tif"))
tmin_04 = raster(paste0(chwd, "tmin/CHELSA_tmin10_04_1979-2013_V1.2_land.tif"))
tmin_05 = raster(paste0(chwd, "tmin/CHELSA_tmin10_05_1979-2013_V1.2_land.tif"))
tmin_06 = raster(paste0(chwd, "tmin/CHELSA_tmin10_06_1979-2013_V1.2_land.tif"))
tmin_07 = raster(paste0(chwd, "tmin/CHELSA_tmin10_07_1979-2013_V1.2_land.tif"))
tmin_08 = raster(paste0(chwd, "tmin/CHELSA_tmin10_08_1979-2013_V1.2_land.tif"))
tmin_09 = raster(paste0(chwd, "tmin/CHELSA_tmin10_09_1979-2013_V1.2_land.tif"))
tmin_10 = raster(paste0(chwd, "tmin/CHELSA_tmin10_10_1979-2013_V1.2_land.tif"))
tmin_11 = raster(paste0(chwd, "tmin/CHELSA_tmin10_11_1979-2013_V1.2_land.tif"))
tmin_12 = raster(paste0(chwd, "tmin/CHELSA_tmin10_12_1979-2013_V1.2_land.tif"))
tmin = stack(tmin_01, tmin_02, tmin_03, tmin_04, tmin_05, tmin_06, tmin_07, tmin_08, tmin_09, tmin_10, tmin_11, tmin_12)
tmin <- crop(tmin, EXT)
tmin <- projectRaster(tmin, crs = prj)
names(tmin) <- c("tmin_01", "tmin_02", "tmin_03", "tmin_04", "tmin_05", "tmin_06", "tmin_07", "tmin_08", "tmin_09", "tmin_10", "tmin_11", "tmin_12")
rm(tmin_01, tmin_02, tmin_03, tmin_04, tmin_05, tmin_06, tmin_07, tmin_08, tmin_09, tmin_10, tmin_11, tmin_12)
tmin_r <- rast(tmin)
tmin_r <- resample(tmin_r, resample_100m, method = "bilinear")
rm(tmin)

# Precipitation amount"Amount" means mass per unit area. "Precipitation" in the
# earth's atmosphere means precipitation of water in all phases.

pr_01 = raster(paste0(chwd, "pr/CHELSA_pr_01_1981-2010_V.2.1.tif"))
pr_02 = raster(paste0(chwd, "pr/CHELSA_pr_02_1981-2010_V.2.1.tif"))
pr_03 = raster(paste0(chwd, "pr/CHELSA_pr_03_1981-2010_V.2.1.tif"))
pr_04 = raster(paste0(chwd, "pr/CHELSA_pr_04_1981-2010_V.2.1.tif"))
pr_05 = raster(paste0(chwd, "pr/CHELSA_pr_05_1981-2010_V.2.1.tif"))
pr_06 = raster(paste0(chwd, "pr/CHELSA_pr_06_1981-2010_V.2.1.tif"))
pr_07 = raster(paste0(chwd, "pr/CHELSA_pr_07_1981-2010_V.2.1.tif"))
pr_08 = raster(paste0(chwd, "pr/CHELSA_pr_08_1981-2010_V.2.1.tif"))
pr_09 = raster(paste0(chwd, "pr/CHELSA_pr_09_1981-2010_V.2.1.tif"))
pr_10 = raster(paste0(chwd, "pr/CHELSA_pr_10_1981-2010_V.2.1.tif"))
pr_11 = raster(paste0(chwd, "pr/CHELSA_pr_11_1981-2010_V.2.1.tif"))
pr_12 = raster(paste0(chwd, "pr/CHELSA_pr_12_1981-2010_V.2.1.tif"))
pr = stack(pr_01, pr_02, pr_03, pr_04, pr_05, pr_06, pr_07, pr_08, pr_09, pr_10, pr_11, pr_12)
pr <- crop(pr, EXT)
pr <- projectRaster(pr, crs = prj)
for (i in 1:nlayers(pr)) {
  pr[[i]] = pr[[i]]/mo[i]
}
names(pr) <- c("pr_01", "pr_02", "pr_03", "pr_04", "pr_05", "pr_06", "pr_07", "pr_08", "pr_09", "pr_10", "pr_11", "pr_12")
rm(pr_01, pr_02, pr_03, pr_04, pr_05, pr_06, pr_07, pr_08, pr_09, pr_10, pr_11, pr_12)
pr_r <- rast(pr)
pr_r <- resample(pr_r, resample_100m, method = "bilinear")
###rm(pr)

# Total % cloud cover for each month
clt_01 = raster(paste0(chwd, "clt/CHELSA_clt_01_1981-2010_V.2.1.tif"))
clt_02 = raster(paste0(chwd, "clt/CHELSA_clt_02_1981-2010_V.2.1.tif"))
clt_03 = raster(paste0(chwd, "clt/CHELSA_clt_03_1981-2010_V.2.1.tif"))
clt_04 = raster(paste0(chwd, "clt/CHELSA_clt_04_1981-2010_V.2.1.tif"))
clt_05 = raster(paste0(chwd, "clt/CHELSA_clt_05_1981-2010_V.2.1.tif"))
clt_06 = raster(paste0(chwd, "clt/CHELSA_clt_06_1981-2010_V.2.1.tif"))
clt_07 = raster(paste0(chwd, "clt/CHELSA_clt_07_1981-2010_V.2.1.tif"))
clt_08 = raster(paste0(chwd, "clt/CHELSA_clt_08_1981-2010_V.2.1.tif"))
clt_09 = raster(paste0(chwd, "clt/CHELSA_clt_09_1981-2010_V.2.1.tif"))
clt_10 = raster(paste0(chwd, "clt/CHELSA_clt_10_1981-2010_V.2.1.tif"))
clt_11 = raster(paste0(chwd, "clt/CHELSA_clt_11_1981-2010_V.2.1.tif"))
clt_12 = raster(paste0(chwd, "clt/CHELSA_clt_12_1981-2010_V.2.1.tif"))
clt = stack(clt_01, clt_02, clt_03, clt_04, clt_05, clt_06, clt_07, clt_08, clt_09, clt_10, clt_11, clt_12)
clt <- crop(clt, EXT)
clt <- projectRaster(clt, crs = prj)
# Resample to new resolution
r2 = raster(extent(pr), resolution = c(875, 922), crs = crs(pr))
clt <- resample(clt, r2, method = "bilinear")
names(clt) <- c("clt_01", "clt_02", "clt_03", "clt_04", "clt_05", "clt_06", "clt_07", "clt_08", "clt_09", "clt_10", "clt_11", "clt_12")
rm(clt_01, clt_02, clt_03, clt_04, clt_05, clt_06, clt_07, clt_08, clt_09, clt_10, clt_11, clt_12, r2)
clt_r <- rast(clt)
clt_r <- resample(clt_r, resample_100m, method = "bilinear")
rm(clt, pr)

# Surface downwelling shortwave flux in air
rsds_01 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_01_V.2.1.tif"))
rsds_02 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_02_V.2.1.tif"))
rsds_03 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_03_V.2.1.tif"))
rsds_04 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_04_V.2.1.tif"))
rsds_05 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_05_V.2.1.tif"))
rsds_06 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_06_V.2.1.tif"))
rsds_07 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_07_V.2.1.tif"))
rsds_08 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_08_V.2.1.tif"))
rsds_09 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_09_V.2.1.tif"))
rsds_10 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_10_V.2.1.tif"))
rsds_11 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_11_V.2.1.tif"))
rsds_12 = raster(paste0(chwd, "rsds/CHELSA_rsds_1981-2010_12_V.2.1.tif"))
rsds = stack(rsds_01, rsds_02, rsds_03, rsds_04, rsds_05, rsds_06, rsds_07, rsds_08, rsds_09, rsds_10, rsds_11, rsds_12)
rsds <- crop(rsds, EXT)
rsds <- projectRaster(rsds, crs = prj)
names(rsds) <- c("rsds_01", "rsds_02", "rsds_03", "rsds_04", "rsds_05", "rsds_06", "rsds_07", "rsds_08", "rsds_09", "rsds_10", "rsds_11", "rsds_12")
rm(rsds_01, rsds_02, rsds_03, rsds_04, rsds_05, rsds_06, rsds_07, rsds_08, rsds_09, rsds_10, rsds_11, rsds_12)
rsds_r <- rast(rsds)
rsds_r <- resample(rsds_r, resample_100m, method = "bilinear")
rm(rsds)


Clim_r <- c(prec_r,pr_r,tmean_r,tmax_r,tmin_r,clt_r,rsds_r)
#names(Clim_r)
writeRaster(Clim_r, filename = paste0(wd$data_p, "Clim_r_resampled.tif"), overwrite=TRUE, append=FALSE)


```



# Arable Area


Streams and their order
Elevation
Slope

0. Load
--DEM
--CONABIO_2015_LandUse_30m
--FASII_Centro_AG_LandUse -- THIS NEEDS TO BE CROPPED
--Mask for Tlaxcala, Hidalgo, Morelos + Some Municipios in DF
--

1. Rasterize 
--SIAP_2015_Fields
--FASII_Centro_AG_LandUse

2. Get all rasters on the same resolution

3. Calculate Topo Vars
--incl streams... geomorphons??
--curv

4. Separate into Training and Prediction
TRAINING
--modern fields in Tlax_Hgo_Mor_partsDF == Arable (Present) == MASK

Subtract modern fields from CONABIO_2015_LandUse_30m


--Sample inarable == Inrable (Absent) **after mask is applied**
Elevation == All areas >= 3100 masl
All Slopes over 30 degrees
All CONABIO_2015_LandUse_30m == 18 or 19
All geomorphons == 2
--Sampled areas from CONABIO_2015_LandUse_30m c(1:14, 16)
--Sampled geomorphons == 9 with higher accum
--Sampled BDRICM < 100
--Sampled BDRLOG > 40
--Sampled Elevation > 2850
--Sampled Slopes > 15

PREDICTION
all areas not in the SIAP_2015_Fields raster

5. Stack all of the data

~~~Predict Irrigation~~~
Both Training and Prediction within Predicted Arable Area
Irrigation sampled from Irrigation land use area (rasterized)

## Prepare Data


### Predictor Variables

```{r, label='Prepare Predictor Data', message=FALSE,warning=FALSE}

### DEM and Topographic Data ###

resample_30m <- rast(paste0(wd$data_r, "resample_30m.tif"))
DEM <- rast(paste0(wd$data_r,"DEM.tif"))
DEM <- resample(DEM, resample_30m, method = "bilinear")
writeRaster(DEM,paste0(wd$data_p,"DEM_30m.tif"), overwrite=TRUE)

TRI <- terra::terrain(DEM, v="TRI")# calc Terrain Ruggedness Index
slope <- terra::terrain(DEM, v="slope", unit="degrees")#Slope
writeRaster(slope,paste0(wd$data_p,"temp_rasts/Slope_30m.tif"), overwrite=TRUE)

wbt_breach_depressions(dem = paste0(wd$data_p,"DEM_30m.tif"), output = paste0(wd$data_p,"temp_rasts/DEM_breach_30m.tif"))

wbt_d8_flow_accumulation(paste0(wd$data_p,"temp_rasts/DEM_breach_30m.tif"), output=paste0(wd$data_p,"temp_rasts/Accum_30m.tif"), out_type = "specific contributing area")

Accum <- rast(paste0(wd$data_p, "temp_rasts/Accum_30m.tif"))
wbt_wetness_index(sca=paste0(wd$data_p,"temp_rasts/Accum_30m.tif"), slope=paste0(wd$data_p,"temp_rasts/Slope_30m.tif"), output=paste0(wd$data_p,"temp_rasts/TWI_30m.tif"), verbose_mode = FALSE)
TWI <- rast(paste0(wd$data_p, "temp_rasts/TWI_30m.tif"))
TWI <- focal(TWI, w=17, fun='max', na.policy="only", na.rm=T)
wbt_stream_power_index(sca=paste0(wd$data_p,"temp_rasts/Accum_30m.tif"),slope=paste0(wd$data_p,"temp_rasts/Slope_30m.tif"),output=paste0(wd$data_p,"temp_rasts/SPI_30m.tif"),exponent = 1)
SPI <- rast(paste0(wd$data_p, "temp_rasts/SPI_30m.tif"))
wbt_fd8_flow_accumulation(paste0(wd$data_p,"temp_rasts/DEM_breach_30m.tif"), output=paste0(wd$data_p,"temp_rasts/Accum_fd8_30m.tif"), out_type = "specific contributing area")
Accum_fd8 <- rast(paste0(wd$data_p, "temp_rasts/Accum_fd8_30m.tif"))
wbt_sediment_transport_index(sca=paste0(wd$data_p,"temp_rasts/Accum_fd8_30m.tif"),slope=paste0(wd$data_p,"temp_rasts/Slope_30m.tif"),output=paste0(wd$data_p,"temp_rasts/STI_30m.tif"))
STI <- rast(paste0(wd$data_p, "temp_rasts/STI_30m.tif"))

Topo_r_30m <- (c(DEM, slope, Accum, Accum_fd8, TRI, TWI, STI, SPI))

nam = c("DEM", "Slope", "Accum", "Accum_fd8", "TRI", "TWI", "STI", "SPI")

names(Topo_r_30m) = nam

rm(slope, Accum, Accum_fd8, TRI, TWI, STI, SPI, nam)

writeRaster(Topo_r_30m, filename = paste0(wd$data_p, "Topo_r_30m.tif"), overwrite=TRUE)

#DEM = Topo_r_15m[[1]]


### Soil Data ###

Soil <- rast(paste0(wd$data_p, "Soil_r_resampled.tif"))

z <- resample(Soil[[1]], DEM, method = "bilinear")

for (i in 2:3) {
  a = resample(Soil[[i]], DEM, method = "bilinear")
  z = c(z,a)
}

Soil_30m = z
rm(z,a)

writeRaster(Soil_30m, filename = paste0(wd$data_p, "Soil_r_30m.tif"), overwrite=TRUE)


### Geomorphons ###


Geomorphons <- rast(paste0(wd$data_r,"CMex_15m_Geomorphons.tif"))
Geomorphons <- as.factor(Geomorphons)
cls = as.data.frame(levels(Geomorphons)[[1]])
cls = cls %>% select(-CMex_15m_Geomorphons) %>%
  mutate(Geomorphons = case_when(
    ID == 1 ~ "Flat",
    ID == 2 ~ "Summit",
    ID == 3 ~ "Ridge",
    ID == 4 ~ "Shoulder",
    ID == 5 ~ "Spur",
    ID == 6 ~ "Slope",
    ID == 7 ~ "Hollow",
    ID == 8 ~ "Footslope",
    ID == 9 ~ "Valley",
    ID == 10 ~ "Depression"))
levels(Geomorphons) <- cls
Geomorphons <- resample(Geomorphons, DEM, method = "near")
writeRaster(Geomorphons, filename = paste0(wd$data_p, "CMex_30m_Geomorphons.tif"), overwrite=TRUE)

CONABIO_2015_LandUse_30m <- rast(paste0(wd$data_r,"CONABIO_2015_LandUse_30m.tif"))
CONABIO_2015_LandUse_30m <- as.factor(CONABIO_2015_LandUse_30m)
cls = levels(CONABIO_2015_LandUse_30m)[[1]] %>% select(-CONABIO_2015_LandUse_30m) %>%
  mutate(LandUse = case_when(
            ID == 1 ~ "Forest_Coniferous_Temperate",
            ID == 2 ~ "Forest_Coniferous_Subpolar",
            ID == 3 ~ "Forest_Evergreen_Subtropical",
            ID == 4 ~ "Forest_Deciduous_Subtropical",
            ID == 5 ~ "Forest_Deciduous_Temperate",
            ID == 6 ~ "Forest_Mixed",
            ID == 7 ~ "Scrub_Subtropical",
            ID == 8 ~ "Scrub_Temperate",
            ID == 9 ~ "Grass_Subtropical",
            ID == 10 ~ "Grass_Temperate",
            ID == 14 ~ "Wetland",
            ID == 15 ~ "Agriculture",
            ID == 16 ~ "Bare_Soil",
            ID == 17 ~ "Settlement",
            ID == 18 ~ "Water",
            ID == 19 ~ "Glacial"),
        LandUseGroups = case_when(
            ID == 1 ~ "Forest_High",
            ID == 2 ~ "Forest_High",
            ID == 3 ~ "Forest_Low",
            ID == 4 ~ "Forest_Low",
            ID == 5 ~ "Forest_Low",
            ID == 6 ~ "Forest_Low",
            ID == 7 ~ "Scrub",
            ID == 8 ~ "Scrub",
            ID == 9 ~ "Grass",
            ID == 10 ~ "Grass",
            ID == 14 ~ "Wetland",
            ID == 15 ~ "Agriculture",
            ID == 16 ~ "Bare_Soil",
            ID == 17 ~ "Settlement",
            ID == 18 ~ "Water",
            ID == 19 ~ "Glacial"),
        Sampling = case_when(
            ID == 1 ~ "Inarable_Sample_High",
            ID == 2 ~ "Inarable_Sample_High",
            ID == 3 ~ "Inarable_Sample_Med",
            ID == 4 ~ "Inarable_Sample_Med",
            ID == 5 ~ "Inarable_Sample_Med",
            ID == 6 ~ "Inarable_Sample_Med",
            ID == 7 ~ "Inarable_Sample_High",
            ID == 8 ~ "Inarable_Sample_High",
            ID == 9 ~ "Inarable_Sample_High",
            ID == 10 ~ "Inarable_Sample_High",
            ID == 14 ~ "Inarable_Sample_Med",
            ID == 15 ~ "Agriculture",
            ID == 16 ~ "Inarable_Sample_High",
            ID == 17 ~ "Settlement",
            ID == 18 ~ "Inarable_All",
            ID == 19 ~ "Inarable_All"),
        Sampling2 = case_when(
            ID == 1 ~ "Sample",
            ID == 2 ~ "Sample",
            ID == 3 ~ "Sample",
            ID == 4 ~ "Sample",
            ID == 5 ~ "Sample",
            ID == 6 ~ "Sample",
            ID == 7 ~ "Sample",
            ID == 8 ~ "Sample",
            ID == 9 ~ "Sample",
            ID == 10 ~ "Sample",
            ID == 14 ~ "Sample",
            ID == 15 ~ "Agriculture",
            ID == 16 ~ "Sample",
            ID == 17 ~ "Settlement",
            ID == 18 ~ "Inarable",
            ID == 19 ~ "Inarable"))

levels(CONABIO_2015_LandUse_30m) <- cls
activeCat(CONABIO_2015_LandUse_30m) <- "Sampling"

CONABIO_2015_LandUse_30m <- resample(CONABIO_2015_LandUse_30m, DEM, method = "near")
CONABIO_2015_LandUse_30m <- crop(CONABIO_2015_LandUse_30m, DEM)

writeRaster(CONABIO_2015_LandUse_30m, filename = paste0(wd$data_p, "CONABIO_2015_LandUse_30m.tif"), overwrite=TRUE)

#Clim_r <- rast(paste0(wd$data_p, "Clim_r_resampled.tif"))
#Clim_r_15m <- subset(Clim_r, c("tmin_04","tmin_05", "tmin_06", "tmin_07", "tmin_08", "tmin_09", "tmin_10","tmin_11"))
#Clim_r_15m <- resample(Clim_r, DEM, method = "bilinear")

```



### Shortcut Import

```{r, label='Shortcut', message=FALSE,warning=FALSE}

DEM <- rast(paste0(wd$data_p, "DEM_30m.tif"))

Topo_r_30m <- rast(paste0(wd$data_p, "Topo_r_30m.tif"))

Soil_30m <- rast(paste0(wd$data_p, "Soil_r_30m.tif"))

Geomorphons <- rast(paste0(wd$data_p, "CMex_30m_Geomorphons.tif"))

CONABIO_2015_LandUse_30m <- rast(paste0(wd$data_p, "CONABIO_2015_LandUse_30m.tif"))

SIAP_2015_Fields <- rast(paste0(wd$data_p, "SIAP_2015_Fields.tif"))
SIAP_2015_Mask <- rast(paste0(wd$data_p, "SIAP_2015_Mask.tif"))
SIAP_2015_Fields <- mask(SIAP_2015_Fields, SIAP_2015_Mask)
SIAP_2015_Fields_f <- as.factor(SIAP_2015_Fields)
levels(SIAP_2015_Fields_f) = levels(SIAP_2015_Fields_f)[[1]] %>% select(-layer) %>%
  mutate(SIAP_2015_Fields = case_when(
    ID == 0 ~ "Uncultivated",
    ID == 1 ~ "Cultivated"))

SIAP_2015_FieldsNA <- ifel(SIAP_2015_Fields == 0, NA, SIAP_2015_Fields)


#Inarable_Sampling_Area <- rast(paste0(wd$data_p, "Inarable_Sampling_Area.tif"))
#InarableSample1 <- rast(paste0(wd$data_p, "InarableSample1.tif"))
#InarableSample2 <- rast(paste0(wd$data_p, "InarableSample2.tif"))
#Inarable_Sampling_Area1 <- rast(paste0(wd$data_p, "Inarable_Sampling_Area1.tif"))
#Inarable_Sampling_Area2 <- rast(paste0(wd$data_p, "Inarable_Sampling_Area2.tif"))

Predictors = c(Topo_r_30m, Soil_30m, Geomorphons, CONABIO_2015_LandUse_30m)
Predictors_FieldsNA <- mask(Predictors, SIAP_2015_FieldsNA)
Predictors = mask(Predictors, SIAP_2015_Mask)
```
```{r}
x = c(SIAP_2015_Fields_f, CONABIO_2015_LandUse_30m, Topo_r_30m, Soil_30m, Geomorphons)
xf <- mask(x, SIAP_2015_FieldsNA)
x = mask(x, SIAP_2015_Mask)





f = freq(x, digits=0, value=NULL, bylayer=TRUE, usenames=T, zones=SIAP_2015_Fields, wide=F)
f = f %>% mutate(zone_f = case_when(
    zone == 0 ~ "Uncultivated",
    zone == 1 ~ "Cultivated"))
c = crosstab(c(SIAP_2015_Fields_f,CONABIO_2015_LandUse_30m), digits=0, long=FALSE, useNA=FALSE)
## S4 method for signature 'SpatRaster,SpatRaster'
z1 = zonal(Predictors, SIAP_2015_Fields, fun="mean", na.rm=T)
z1$stat = "Mean"
z2 = zonal(Predictors, SIAP_2015_Fields, fun="min", na.rm=T)
z2$stat = "Min"
z3 = zonal(Predictors, SIAP_2015_Fields, fun="max", na.rm=T)
z3$stat = "Max"
z = rbind(z1, z2, z3) %>% mutate(zone = case_when(
    layer == 0 ~ "Uncultivated",
    layer == 1 ~ "Cultivated"))

x_df <- as.data.frame(x[[1]])
names(x_df) <- "SIAP_2015_Fields_f"
x_df <- rownames_to_column(x_df, var = "Cells")

for (i in 2:nlyr(x)) {
  
  t = as.data.frame(x[[i]])
  t <- rownames_to_column(t, var = "Cells")
  names(t) <- c("Cells", names(x[[i]]))
  x_df <- x_df %>% left_join(t, by="Cells")
  #Train_df <- cbind(Train_df, t[,1])
  #names(Train_df) <- c(names(Train_df)[1:(i-1)], names(Train[[i]]))
}
x_df = x_df[complete.cases(x_df), ]



write_csv(x_df, paste0(wd$data_p,"ArableArea_RFModel_TrainData.csv.gz"))
Arable_RF_Train_df <- read_csv(paste0(wd$data_p,"ArableArea_RFModel_TrainData.csv.gz"))

rm(x_df)


selected_rows <- Arable_RF_Train_df %>%
  filter(var1 > 3 & var2 == "A") %>%
  slice_sample(n = n, replace = FALSE)


LandUse == "Cultivated"




```
write.csv(x_df, paste0(wd$data_r,"ArableArea_RFModel_TrainData.csv"))
library(readr)
write_csv(x_df, paste0(wd$data_p,"ArableArea_RFModel_TrainData.csv.gz"))
Arable_RF_Train_df <- read_csv(paste0(wd$data_p,"ArableArea_RFModel_TrainData.csv.gz"))
TRI > 26

25.938477

BDRICM < 40


CROSSTAB Cross-tabulate the layers of a SpatRaster to create a contingency table.
**x**; SpatRaster
**digits**; integer. The number of digits for rounding the values before cross-tabulation
**long**; logical. If TRUE the results are returned in ’long’ format data.frame instead of a table
**useNA**; logical, indicting if the table should includes counts of NA values

FREQ Frequency table of the values of a SpatRaster. NAs are not counted unless value=NA.**You can provide a SpatVector or additional SpatRaster to define zones for which to do tabulations.**
**x**; SpatRaster
**digits**; integer. Used for rounding the values before tabulation. Ignored if NA
**value**; numeric. An optional single value to only count the number of cells with that value. This value can be NA
**bylayer**; logical. If TRUE tabulation is done by layer
**usenames**; logical. If TRUE layers are identified by their names instead of their numbers.
Only relevant if bylayer is TRUE
**zones**; SpatRaster or SpatVector to define zones for which the tabulation should be done
**wide**; logical. Should the results by "wide" instead of "long"?

zonal Zonal statistics
Compute zonal statistics, that is summarize values of a SpatRaster for each "zone" defined by another SpatRaster, or by a SpatVector with polygon geometry.
If fun is a true R function, the <SpatRaster,SpatRaster> method may fail when using very large
SpatRasters, except for the functions ("mean", "min", "max", "sum", "isNA", and "notNA").
You can also summarize values of a SpatVector for each polygon (zone) defined by another SpatVec

x SpatRaster or SpatVector
z SpatRaster with cell-values representing zones or a SpatVector with each polygon geometry representing a zone. z can have multiple layers to define intersecting zones
fun function to be applied to summarize the values by zone. Either as character:
"mean", "min", "max", "sum", "isNA", and "notNA" and, for relatively small
SpatRasters, a proper function
... additional arguments passed to fun, such as na.rm=TRUE
w SpatRaster with weights. Should have a single-layer with non-negative values
wide logical. Should the values returned in a wide format? This only affects the
results when nlyr(z) == 2
as.raster logical. If TRUE, a SpatRaster is returned with the zonal statistic for each zone
filename character. Output filename (ignored if as.raster=FALSE
overwrite logical. If TRUE, filename is overwritten
wopt list with additional arguments for writing files as in writeRaster
weights logical. If TRUE and y has polygons, the approximate fraction of each cell that is
covered is returned as well, for example to compute a weighted mean
### Dependent Variable

#### Import and Rasterize Data

```{r, label='Shortcut', message=FALSE,warning=FALSE}
SIAP_2015_Fields <- vect(paste0(wd$data_r,"SIAP_2015_Fields.gpkg"))
SIAP_2015_Fields <- rasterize(SIAP_2015_Fields, DEM, values = 1, background = 0)
writeRaster(SIAP_2015_Fields, filename = paste0(wd$data_p, "SIAP_2015_Fields.tif"), overwrite=TRUE)

SIAP_2015_Mask <- vect(paste0(wd$data_r,"SIAP_2015_Mask.gpkg"))
SIAP_2015_Mask <- terra::rasterize(SIAP_2015_Mask, DEM, values = 1, background = NA)
writeRaster(SIAP_2015_Mask, filename = paste0(wd$data_p, "SIAP_2015_Mask.tif"), overwrite=TRUE)

SIAP_2015_Fields <- mask(SIAP_2015_Fields, SIAP_2015_Mask)

#rc <- c(-0.5, 0.5, 1,  0.5, 1.5, NA)
#rc <- matrix(c(-0.5, 0.5, 1,  0.5, 1.5, NA), ncol=3, byrow=TRUE)
Inarable_Sampling_Area <- classify(SIAP_2015_Fields, matrix(c(-0.5, 0.5, 1,  0.5, 1.5, NA), ncol=3, byrow=TRUE))
writeRaster(Inarable_Sampling_Area, filename = paste0(wd$data_p, "Inarable_Sampling_Area.tif"), overwrite=TRUE)

rm(SIAP_2015_Fields)

```


#### Absence (Inarable) Sample of Background Points



##### Inarable Sample #1: Totally Inarable

```{r, label='Inarable Sample #1: Totally Inarable', message=FALSE,warning=FALSE}

InarableSample1 = Inarable_Sampling_Area

#Elevation >= 3100 masl
q = DEM * InarableSample1
q = ifel(q >= 3100, 1, 0)
InarableSample1 <- q + InarableSample1

#CONABIO_2015_LandUse_15m == 18 and 19
q = ifel(CONABIO_2015_LandUse_30m == 18 | CONABIO_2015_LandUse_30m == 19, 1, 0)
InarableSample1 <- q + InarableSample1

InarableSample1 <- ifel(InarableSample1 > 1, 1, NA)
Inarable_Sampling_Area1 <- InarableSample1
writeRaster(Inarable_Sampling_Area1, filename = paste0(wd$data_p, "Inarable_Sampling_Area1.tif"), overwrite=TRUE)

#table(values(Sampling_Area1))
s = spatSample(InarableSample1, size = 50000, method="random", replace=FALSE, na.rm=T, exhaustive=T, cells=T) #40000 = 0.0639091

v <- unique(s$cell)
InarableSample1 = ifel(InarableSample1 > 0, 0, NA)
InarableSample1[v] <- 2
InarableSample1 = ifel(InarableSample1 > 1, 2, 0)

writeRaster(InarableSample1, filename = paste0(wd$data_p, "InarableSample1.tif"), overwrite=TRUE)

InarableSample1 <- rast(paste0(wd$data_p, "InarableSample1.tif"))

rm(s, v, q)

```


##### Inarable Sample #2: High Likelihood Inarable

```{r, label='Inarable Sample #2: High Likelihood Inarable', message=FALSE,warning=FALSE}

### Sampling Area2
InarableSample2 <- mosaic(Inarable_Sampling_Area1, Inarable_Sampling_Area, fun=sum)
InarableSample2 = ifel(InarableSample2 > 1, NA, 0)

#Sampled areas from CONABIO_2015_LandUse_15m c(1:14, 16)
q = ifel(CONABIO_2015_LandUse_30m < 15 | CONABIO_2015_LandUse_30m == 16, 1, 0)
InarableSample2 <- q + InarableSample2

#Sampled geomorphons == 9,10 with higher accum
q = ifel(Geomorphons == 9 | Geomorphons == 10 | Geomorphons == 2, 1, 0)
InarableSample2 <- q + InarableSample2

#Sampled BDRICM < 100
q = ifel(Soil_30m[["BDRICM"]] < 100, 1, 0)
InarableSample2 <- q + InarableSample2

#Sampled BDRLOG > 40
q = ifel(Soil_30m[["BDRLOG"]] > 40, 1, 0)
InarableSample2 <- q + InarableSample2

#Sampled Elevation > 2850
q = ifel(DEM >= 2850, 1, 0)
InarableSample2 <- q + InarableSample2

#Sampled Slopes > 15
q = ifel(Topo_r_30m[["Slope"]] > 15, 1, 0)
InarableSample2 <- q + InarableSample2

#STI >15*********
q = ifel(Topo_r_30m[["STI"]] > 15, 1, 0)
InarableSample2 <- q + InarableSample2

#TRI >15
q = ifel(Topo_r_30m[["TRI"]] > 15, 1, 0)
InarableSample2 <- q + InarableSample2
#Accum>4500
#TWI > 15
#SPI >300-500

t = freq(InarableSample2)
t = t[-c(1),]
t$prob = c(0.004, 0.008, 0.015, 0.03, 0.06, 0.25, 1.0)
t$sample = round((t$prob * t$count),0)
sum(t$sample)

# Create a list to store the individual rasters
r_list <- list()

# Loop through each unique value
for (i in t$value) {
  # Create a binary mask for the current value
  masked_raster <- ifel(InarableSample2 == t$value[i], 1, NA)
  
  s2 = spatSample(x = masked_raster, size = t$n[i], method="random", replace=FALSE, na.rm=T, exhaustive=T, cells=T) #

  # Store the masked raster in the list
  r_list[[as.character(unique_values[i])]] <- s2
}

s2 = do.call(rbind, r_list)

v <- unique(s2$cell)
InarableSample2 <- ifel(Inarable_Sampling_Area > 0, 0, NA)
InarableSample2[v] <- 2
InarableSample2 = ifel(InarableSample2 > 1, 2, 0)

writeRaster(InarableSample2, filename = paste0(wd$data_r, "InarableSample2.tif"), overwrite=TRUE)

rm(s, v, q, s2, t)

```
##### Combining the Inarable Samples

```{r}

InarableSample <- mosaic(inarable1, inarable2, fun=sum)
writeRaster(InarableSample, filename = paste0(wd$data_r, "InarableSample.tif"), overwrite=TRUE)
```


##### Arable Sample

```{r}
Arable <- rasterize(SIAP_2015_Fields, DEM, values = 0, background = NA) #values = 0
table(values(Arable))
sa = spatSample(Arable, size = 233000, method="random", replace=FALSE, na.rm=T, exhaustive=T, cells=T) #225000 == 0.06222285

v <- unique(sa$cell)
ArableSamp = Arable
ArableSamp[v] <- 2
ArableSamp = ifel(ArableSamp > 1, 2, 0)
ArableSample <- mosaic(Arable, ArableSamp, fun=sum)

writeRaster(ArableSample, filename = paste0(wd$data_r, "ArableSample.tif"), overwrite=TRUE)


```



### Training Dataset


```{r}
rm(q, s, s2, sa, Sampling_Area2, Sampling_Area1, r_list, v, mask, masked_raster, cls, rc, t, wts, Soil, inarable1, inarable2, ArableSamp, Arable)

Covariates = c(Topo_r_30m, Soil_30m, Geomorphons, CONABIO_2015_LandUse_30m)

rm(Topo_r_30m, Soil_30m, Geomorphons, CONABIO_2015_LandUse_30m)

Prediction2 <- ifel(InarableSample == 0, 1, NA)
Predict_Supp_ArableUnsampled <- ifel(ArableSample == 0, 1, NA)#ArableSample == 0
InarableSample = ifel(InarableSample > 0, 1, 0)
InarableSample = ifel(is.na(InarableSample), 0, InarableSample)
ArableSample = ifel(is.na(ArableSample), 0, ArableSample)
Response <- mosaic(ArableSample, InarableSample, fun=sum)
Response <- mask(Response, SIAP_2015_Mask)
Prediction <- ifel(Response == 0, 1, NA)
Response <- ifel(Response == 0, NA, Response)

Response <- as.factor(Response)
cls = as.data.frame(levels(Response)[[1]])
cls = cls %>% select(-layer) %>%
  mutate(Response = case_when(
    ID == 1 ~ "Inarable",
    ID == 2 ~ "Arable"))
levels(Response) <- cls

Train = c(Response, Covariates)
Train <- mask(Train, Response)



Predict = c(Prediction, Covariates)
Predict <- mask(Predict, Prediction)

#plot(Train[[2]])
#plot(Predict[[2]])

rm(SIAP_2015_Fields, ArableSample, InarableSample, cls, DEM)



activeCat(Train[[14]]) <- "LandUseGroups"

Train_df <- as.data.frame(Train[[1]])
names(Train_df) <- "Arable"
Train_df <- rownames_to_column(Train_df, var = "Cells")

for (i in 2:nlyr(Train)) {
  
  t = as.data.frame(Train[[i]])
  t <- rownames_to_column(t, var = "Cells")
  names(t) <- c("Cells", names(Train[[i]]))
  Train_df <- Train_df %>% left_join(t, by="Cells")
  #Train_df <- cbind(Train_df, t[,1])
  #names(Train_df) <- c(names(Train_df)[1:(i-1)], names(Train[[i]]))
}
Train_df = Train_df[complete.cases(Train_df), ]

```


### Prediction Dataset

```{r}

activeCat(Predict[[14]]) <- "LandUseGroups"

Predict_df <- as.data.frame(Predict[[2]])
names(Predict_df) <- "DEM"
Predict_df <- rownames_to_column(Predict_df, var = "Cells")

for (i in 3:nlyr(Predict)) {
  
  t = as.data.frame(Predict[[i]])
  t <- rownames_to_column(t, var = "Cells")
  names(t) <- c("Cells", names(Predict[[i]]))
  Predict_df <- Predict_df %>% left_join(t, by="Cells")

}
Predict_df = Predict_df[complete.cases(Predict_df), ]

```


## Random Forest Model

### Variable Selection

```{r}


boruta <- Boruta(Arable ~ ., data = Train_df, maxRuns = 1000)#, doTrace = 2
print(boruta)
boruta[["finalDecision"]]
plot(boruta, las = 2, cex.axis = 0.7)
#bor <- TentativeRoughFix(boruta)
#bor[["finalDecision"]]
attStats(boruta)
form=getConfirmedFormula(boruta)
```


### Train the Model

```{r}

RF_model <- ranger(Arable ~ DEM + Slope + Accum_fd8 + TRI + TWI + STI + SPI + BDRICM + BDRLOG + BDTICM + Geomorphons + LandUseGroups, data = Train_df, importance = 'impurity')

print(RF_model)

```

### Model Predictions

```{r}

RF_pred <- predict(RF_model, data=Predict_df)

```

  predicted_vals <- predict(model, pred)
  pred$pred <- predicted_vals$predictions
  
  x = r_df %>% left_join(pred, by = f)
  
  r2 = r
  raster_values <- values(r2)
  missing_values <- is.na(raster_values)
  df_values <- x$pred[missing_values]
  r2[missing_values] <- df_values
  names(r2) <- rnam
  
### Model Cross-Validation

```{r}
p.ra <- predict(m.lzn.ra, data=meuse)
str(p.ra)

summary(r.rap <- meuse$logZn - p.ra$predictions)

(rmse.ra <- sqrt(sum(r.rap^2)/length(r.rap)))

plot(meuse$logZn ~ p.ra$predictions, asp=1, pch=20, xlab="fitted", ylab="actual", xlim=c(2,3.3),          ylim=c(2,3.3), main="log10(Zn), Meuse topsoils, Ranger")
grid(); abline(0,1)


summary(m.lzn.ra$predictions)

plot(meuse$logZn ~ m.lzn.ra$predictions, asp=1, pch=20,
     ylab="actual", xlab="OOB X-validation estimates",
    xlim=c(2,3.3), ylim=c(2,3.3),
    main="ranger")
abline(0,1); grid()

```
















# Extras

```{r}

#rc <- c(-101, -98, 1,  0.25, 1.5, NA)
#rc <- matrix(rc, ncol=3, byrow=TRUE)
#Sampling_Area <- classify(SIAP_2015_Fields, rc)
                         
f <- names(train)
  f <- f[f != names(r)]
  form <- paste(names(r), "~", paste(f, collapse = " + "))
  form <- as.formula(form)
  

```


mtry = NULL,

probability = T #Grow a probability forest as in Malley et al. (2012).


regularization.factor = 1,Regularization factor (gain penalization), either a vector of length p or one value
for all variables.

regularization.usedepth = FALSE,Consider the depth in regularization

quantreg = FALSE,Prepare quantile prediction as in quantile regression forests (Meinshausen 2006).
Regression only. Set keep.inbag = TRUE to prepare out-of-bag quantile prediction.

importance = "none",Variable importance mode, one of ’none’, ’impurity’, ’impurity_corrected’, ’permutation’. The ’impurity’ measure is the Gini index for classification, the variance of the responses for regression and the sum of test statistics (see splitrule)
for survival.
keep.inbag = FALSE,
inbag = NULL,Save how often observations are in-bag in each tree.

save.memory = FALSE  Use memory saving (but slower) splitting mode. No effect for survival and
GWAS data. Warning: This option slows down the tree growing, use only if you
encounter memory problems.

classification = NULL, Set to TRUE to grow a classification forest. Only needed if the data is a matrix or
the response numeric.

midpoints of all arable polygons;
negative buffer so observations arent on edges



OOB (Out-of-Bag) prediction error, also known as OOB error, is a measure of prediction accuracy for models trained using bootstrap aggregating (bagging) techniques, such as random forests. It is commonly used to estimate the performance of a model without the need for an explicit validation set.

When training a bagging model, each base model (also known as a weak learner) is trained on a random subset of the original data, called a bootstrap sample. The remaining data points, on average around 36.8% of the original data, are not included in the bootstrap sample and are referred to as out-of-bag (OOB) observations.

The OOB prediction error is calculated by evaluating the performance of each base model on the OOB observations. For each data point that was not included in the bootstrap sample for a particular base model, the model's prediction is compared to the true target value. The OOB error is then computed as the average prediction error across all out-of-bag observations.

The OOB error serves as an estimate of the model's performance on unseen data. It can be useful for assessing the generalization capability of the model and comparing different models or tuning parameters within the same algorithm. However, it's important to note that the OOB error might not be as reliable as a dedicated validation set for model evaluation in some cases.

In summary, the OOB prediction error provides an estimate of the model's performance using the observations that were not included in the training process due to the bootstrap sampling. It is a useful tool for evaluating bagging-based models such as random forests.

